---
title: 'Model inference Q1'
author: "Erick Calderon-Morales"
date: ''
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: tactile
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r, setup, include=FALSE}
# Set knit directory
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,

                      # Save figures as pdf ?
                      # dev = c( "png", "pdf"),

                      # Include code?
                      echo           = FALSE,

                      error          = FALSE,
                      fig.align      = "center",

                      # Path where figures are going to be store pdf single
                      # figures
                      fig.path       = paste0("./figures_inference", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r cleanup-docs, cache = FALSE,echo = FALSE}
# save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#     file.rename(file, file.path("../../[insert_folder_name]/", file))
# }
```


```{r libaries, message = FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For tables appearance
library(reactablefmtr)
# For cleaning colnames
library(janitor)
# For add column function
library(tibble)
library(purrr)
library(performance)
# For bootstrap confidence intervals
library(lmeresampler)
library(nlme)
library(lme4)
# For model comparison
library(MuMIn)
# For get preditions
library(modelr)
# For reporting results
library(report)
library(emmeans)
```

# Load functions and data

```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
source("./scripts/code_join_data_full_dataset.R")
```

```{r message=FALSE, warning=FALSE}
# Load functions

## Inference
source("./R/function_for_inference_anova_table.R")
source("./R/function_for_inference_tukey_tables.R")
source("./R/function_for_inference_emmeans_and_percentage_diff.R")

## Plots
source("./R/function_masomenos_plots.R")
```


__Q2: How does increased nutrient and/or water availability influence seedling
water- and nutrient-use traits and the relationships with N-fixing bacteria?__

# Load models

```{r}
models_q2 <- readRDS("./processed_data/models_q2.RData")
model_nodule_count <- readRDS("./processed_data/model_nodule_log.RData")
```


```{r}
names(models_q2)
```

```{r echo=FALSE}
# Step done for getting predictions from models for Q2
data_for_models <-
    data_for_models %>%

        rownames_to_column("id") %>%

        # Remove unused variables
        dplyr::select(id, spcode, treatment, nfixer, init_height)
```

```{r}
# Step done for getting predictions from nodule models
source("./scripts/code_clean_data_nodules.R")

data_pred_nodules <-
     data_nodules_cleaned %>%

         # add id to rownames for keep track of the rows
         dplyr::select(c("id", "spcode", "treatment", "init_height"))

```

# Traits

## ANOVAs

```{r}
models_q2 %>%

    # Remove traits and mass fractions models
    anova_table_tidy(., model_list = T)
```

## Tuckey comparisons

### Models with significant interaction

```{r}
tukey_table_tidy(models_q2$amax, formula = "treatment|nfixer")
```

```{r}
tukey_table_tidy(models_q2$wue_log, formula = "treatment|nfixer")
```

### Models with signifcant treatment effect

```{r}
tukey_table_tidy(models_q2$gs, formula = "treatment")
```

```{r}
tukey_table_tidy(models_q2$pnue_log, formula = "treatment")
```

```{r}
tukey_table_tidy(models_q2$nitrogen_log, formula = "treatment")
```


### Models with signifcant nfixer effect

```{r}
tukey_table_tidy(models_q2$gs, formula = "nfixer")
```

```{r}
tukey_table_tidy(models_q2$nitrogen_log, formula = "nfixer")
```

## Percentange change

### Models with significant interaction

```{r}
emmeans_table_tidy(models_q2$amax, formula = "treatment|nfixer",
                    grouping_var = "nfixer")
```


```{r}
emmeans_table_tidy(models_q2$wue_log, formula = "treatment|nfixer",
                    grouping_var = "nfixer")
```


### Models with significant treatment effect

```{r}
emmeans_table_tidy(models_q2$gs, formula = "treatment")
```

```{r}
emmeans_table_tidy(models_q2$pnue_log, formula = "treatment")
```

```{r}
emmeans_table_tidy(models_q2$nitrogen_log, formula = "treatment")
```


### Models with signifcant nfixer effect

```{r}

# Difference in gs between fixer non fixers
# Percentage difference
as.data.frame(emmeans(models_q1_q2$gs,
                          specs = pairwise ~ "nfixer",
                          type = "response",
                          adjust = "tukey")$emmeans) %>%

        clean_names() %>%
        dplyr::select(-c(df,lower_cl, upper_cl,se)) %>%

        # Rename response to emmean, this is done when models is log


        mutate(difference = ((emmean - first(emmean))),
               perc_difference =((emmean - first(emmean) )/first(emmean))*100) %>%
        mutate_if(is.numeric, round, 3)
```


```{r}
as.data.frame(emmeans(models_q1_q2$nitrogen_log,
                          specs = pairwise ~ "nfixer",
                          type = "response",
                          adjust = "tukey")$emmeans)  %>%

        clean_names() %>%
        dplyr::select(-c(df, lower_cl, upper_cl, se)) %>%

        # Rename response to emmean, this is done when models is log


        mutate(difference = ((response - first(response))),
               perc_difference =((response - first(response) )/first(response))*100) %>%
        mutate_if(is.numeric, round, 3)

```


## Plot Traits boxplot

```{r}
string <- c("models_q2")

data_pred_traits <-

        # Get models prediction
        gather_predictions(data_for_models,

                           # Return predictions
                            models_q2$amax,
                            #models_q2$d13c,
                            models_q2$gs,
                            models_q2$wue_log,
                            models_q2$pnue_log) %>%



        pivot_wider(names_from = model, values_from = pred) %>%
            rename_all(funs(

                # rename columns
                stringr::str_to_lower(.) %>%
                stringr::str_replace(., c(string),"pred_") %>%

                # Remove dollar sing
                gsub("\\$", "", .)
                )) %>%

        # Back transform log variables
        mutate(pred_wue = exp(pred_wue_log),
               pred_pnue = exp(pred_pnue_log)) %>%

        # Remove log predictions and init height
        dplyr::select(-c(init_height, pred_wue_log, pred_pnue_log))
```


```{r}
vars_q2 <-
  crossing(

    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_traits, where(is.numeric)) %>% names, 1))),

    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_traits, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_traits, treatment) %>%  names)
```

```{r fig_2_traits_q2_boxplot, fig.height = 6, fig.width = 7, fig.align = 'center'}
vars_q2 %>%
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_traits,
                                y = !!sym(..1), x = !!sym(..2),
                                fill = !!sym(..3)))
#      cowplot::plot_grid(plotlist = ., ncol = 2)

```




```{r}
models_q2 %>%
    map(., r2) %>%
    unlist()
```


# Nodule Count

```{r}
# Interpretation: https://stats.oarc.ucla.edu/other/mult-pkg/faq/general/faqhow-do-i-interpret-a-regression-model-when-some-variables-are-log-transformed/
```


## ANOVAs

```{r}
anova.lme(model_nodule_count, type = "marginal")
```

## Tuckey comparisons

```{r}
as.data.frame(emmeans(model_nodule_count,
                           specs = pairwise ~ "treatment",
                           type = "response",
                           adjust =  "tukey")$contrast) %>%
         clean_names() %>%
         reactable()
```

## Percentage of change

```{r}
# Percentage difference
emmeans_table_tidy(model_nodule_count,
                        formula = "treatment")
```