---
title: 'Model inference Q1'
author: "Erick Calderon-Morales"
date: ''
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: tactile
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r, setup, include=FALSE}
# Set knit directory
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r knitr, include = FALSE}

# Save figures in specific place
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      comment = "",

                      # Save figures as pdf ?
                      dev = c( "png", "pdf"),

                      # Include code?
                      echo           = TRUE,

                      error          = FALSE,
                      fig.align      = "center",

                      # Path where figures are going to be store pdf single
                      # figures
                      fig.path       = paste0("./figures_inference", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)

# set to two decimal
options(scipen = 1, digits = 4)
```

```{r cleanup-docs, cache = FALSE,echo = FALSE}
# save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#     file.rename(file, file.path("../../[insert_folder_name]/", file))
# }
```


```{r libaries, message = FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For r2
library(performance)
# For bootstrap confidence intervals
library(nlme)
library(lme4)
# For model comparison
library(MuMIn)
# For get preditions
library(modelr)
# For rendering tables
library(reactablefmtr)
library(ggplot2)
# For effect sizes
library(effectsize)
library(emmeans)
```

# Load functions and data

```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
source("./scripts/code_join_data_full_dataset.R")
```

```{r message=FALSE, warning=FALSE}
# Load functions

## Inference
source("./R/function_for_inference_anova_table.R")
source("./R/function_for_inference_tukey_tables.R")
source("./R/function_for_inference_emmeans_and_percentage_diff.R")

## Plots
source("./R/function_masomenos_plots.R")
```


__Q1: What is the relative influence of water and/or nutrient availability on
tropical dry forest seedling growth and biomass allocation__


# Load models

```{r}
models_q1 <- readRDS("./processed_data/models_q1.RData")
```


```{r}
# Model available
names(models_q1)
```

```{r echo=FALSE}

# Step done for getting predictions from models for Q1
data_for_predictions <-
    data_for_models %>%

        rownames_to_column("id") %>%

        # Remove unused variables
        dplyr::select(id, spcode, treatment, nfixer, init_height)
```


# ANOVAs

All models show a significant two-way interaction between Nfixer and Treatment

```{r}
models_q1 %>%

    # Remove traits and mass fractions models
    purrr::list_modify("agr_log" = NULL, "agr" = NULL) %>%
    anova_table_tidy(., model_list = T)

```

# Post-Hoc

The main argument of the paper is:

_Results from our shade house experiment show that nutrient addition had a
stronger effect on tropical dry forest seedling biomass and growth than water
addition. Moreover, we demonstrate that N-fixing species have a higher capacity
to take advantage of increased resource availability, particularly for nutrients,
and to a lesser degree water..._


__however I think this is not true since there are not significant differences
when the same treatment are compared (i.e. plus nutrients nfixer vrs plus nutrients non-nfixer)__

## Total Biomass


### How I reported in the paper

```{r}
as_data_frame(emmeans(models_q1$total_biomass,
        pairwise ~ treatment|nfixer,
        adjust = "tukey"
        )$contrast)  %>%
        select(-c(df, SE)) %>%
        reactable()
```


### How it should be done

```{r}
as_data_frame(emmeans(models_q1$total_biomass,
        pairwise ~ treatment*nfixer,
        adjust = "tukey"
        )$contrast)  %>%
        select(-c(df, SE)) %>%
        reactable()

```


### Percentage of Change
```{r}
emmeans_table_tidy(models_q1$total_biomass,
                        formula = "treatment|nfixer",
                        grouping_var = "nfixer")
```


## RGR


### How I reported in the paper

```{r}
as_data_frame(emmeans(models_q1$rgr,
        pairwise ~ treatment|nfixer,
        adjust = "tukey"
        )$contrast)  %>%
        select(-c(df, SE)) %>%
        reactable()

```


### How it should be done

```{r}
as_data_frame(emmeans(models_q1$rgr,
        pairwise ~ treatment*nfixer,
        adjust = "tukey"
        )$contrast) %>%
        select(-c(df, SE)) %>%
        reactable()
```


### Percentage of Change

```{r}
emmeans_table_tidy(models_q1$rgr,
                        formula = "treatment|nfixer",
                        grouping_var = "nfixer")
```



## Root shoot ratio

### How I reported in the paper

```{r}
as_data_frame(emmeans(models_q1$root_shoot_ratio_log,
        pairwise ~ treatment|nfixer,
        adjust = "tukey"
        )$contrast)  %>%
        select(-c(df, SE)) %>%
        reactable()
```


### How it should be done

```{r}
as_data_frame(emmeans(models_q1$root_shoot_ratio_log,
        pairwise ~ treatment*nfixer,
        adjust = "tukey"
        )$contrast)  %>%
        select(-c(df, SE)) %>%
        reactable()
```


## Percentage of Change

```{r}
emmeans_table_tidy(models_q1$root_shoot_ratio_log,
                        formula = "treatment|nfixer",
                        grouping_var = "nfixer")
```


# Plots Biomass and growth boxplot

```{r message=FALSE, warning=FALSE}

# Get predictions
string <- c("models_q1")

data_pred_biomass_growth <-

        # Get models prediction
        gather_predictions(data_for_predictions,

                           # Return predictions of:
                            models_q1$rgr,
                            models_q1$total_biomass,
                            models_q1$root_shoot_ratio_log)  %>%

            # Get fitted values
            pivot_wider(names_from = model, values_from = pred) %>%
             rename_all(
              funs(

                # rename columns
                stringr::str_to_lower(.) %>%

                # Remove string from name and replace it with pred_
                stringr::str_replace(., c(string), "pred_") %>%

                # Remove dollar sing
                gsub("\\$", "", .))) %>%


        # Back transform log variables
        mutate(pred_root_shoot_ratio = exp(pred_root_shoot_ratio_log)) %>%

        # Select only pred variables
        dplyr::select(-c(init_height, pred_root_shoot_ratio_log))
```

```{r warning=FALSE}

# Generate plot combinations
vars_q1 <-
  crossing(

    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_biomass_growth, where(is.numeric)) %>% names, 1))),

    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_biomass_growth, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_biomass_growth, treatment) %>%  names)
```

```{r fig_1_biomass_growth_boxplot, fig.align = 'center', out.width=1000, out.height=900, dpi=350}
par(mfrow = c(2,2))
vars_q1 %>%
      # Gererate plots
      pmap( ~boxplot_plot_pmap(data = data_pred_biomass_growth,
                                y = !!sym(..1), x = !!sym(..2),
                                fill = !!sym(..3)))
```

## r2 models

```{r}
models_q1 %>%

    # Remove traits and mass fractions models
    purrr::list_modify("agr_log" = NULL, "agr" = NULL) %>%

    map(., r2) %>%
    unlist()
```

