---
title: 'Model inference Q3'
author: "Erick Calderon-Morales"
date: ''
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: tactile
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r, setup, include=FALSE}
# Set knit directory
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      comment = "",

                      # Save figures as pdf ?
                      # dev = c( "png", "pdf"),

                      # Include code?
                      echo           = TRUE,

                      error          = FALSE,
                      fig.align      = "center",

                      # Path where figures are going to be store pdf single
                      # figures
                      fig.path       = paste0("./figures_inference", "/"),
                      #fig.width      = 11,
                      #fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r cleanup-docs, cache = FALSE,echo = FALSE}
# save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#     file.rename(file, file.path("../../[insert_folder_name]/", file))
# }
```


```{r libaries, message = FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For tables appearance
library(reactable)
# For cleaning colnames
library(janitor)
# For add column function
library(tibble)
#library(purrr)
#library(performance)
# For bootstrap confidence intervals
library(lmeresampler)
library(nlme)
#library(lme4)
# For model comparison
library(MuMIn)
# For get preditions
#library(modelr)
# For reporting results
library(report)
library(ggplot2)
```

# Load functions and data

```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
source("./scripts/code_join_data_full_dataset.R")
```

```{r message=FALSE, warning=FALSE}
# Load functions for Inference

## Bootstrap
source("./R/function_for_inference_bootstrap.R")


## Plots
source("./R/function_masomenos_plots.R")
```


__Q3: How do seedling-level functional traits, treatment, and N-fixing status
interact to influence seedling-level growth rates and biomass accumulation?__


# Load data

```{r}
data_for_models <-
    data_for_models %>%

    # Select variables for analysis
    dplyr::select(!c(
        rgr_slope, d15n, above_biomass, below_biomass,
        rmf, smf, lmf))
```


## Scale preditors aka traits

```{r}
data_for_models_scaled <-
    data_for_models  %>%

    #  Narea_g_m2 not included since it has a high correlation with amax
    mutate(across(c("init_height", "amax", "gs", "wue", "pnue"), scale))
```


# RGR

```{r}
lme_rgr_3way <- lme(rgr ~   treatment:nfixer:amax +
                            treatment:nfixer:gs  +
                            treatment:nfixer:wue +
                            treatment:nfixer:pnue +
                            #treatment:nfixer:d13c +

                            # Added
                            #treatment:nfixer:Narea_g_m2 +
                            init_height ,
                  random = ~1 | spcode,
                  data = data_for_models_scaled)
```

## Model Selection

```{r}
dredge(lme_rgr_3way)
```


## R2
```{r}
r.squaredGLMM(lme_rgr_3way)
```


# AGR

```{r}
lme_agr_3way <- lme(agr ~   treatment:nfixer:amax +
                            treatment:nfixer:gs  +
                            treatment:nfixer:wue +
                            treatment:nfixer:pnue +
                            #treatment:nfixer:d13c +

                            # Added
                            #treatment:nfixer:Narea_g_m2 +
                            init_height ,
                  random = ~1 | spcode,
                  data = data_for_models_scaled)
```

## Model Selection

```{r}
dredge(lme_agr_3way)
```

## R2

```{r}
r.squaredGLMM(lme_agr_3way)
```


# Total Biomass

```{r}
lme_total_biom_3way <- lme(total_biomass ~  treatment:nfixer:amax +
                                            treatment:nfixer:gs  +
                                            treatment:nfixer:wue +
                                            treatment:nfixer:pnue +
                                            #treatment:nfixer:d13c +

                                            init_height,
                            random = ~1 | spcode,
                            data = data_for_models_scaled)
```


## Model Selection

```{r}
dredge(lme_total_biom_3way)
```

## R2

```{r}
r.squaredGLMM(lme_total_biom_3way)
```


```{r}
# Look results before bootstrap
report::report(lme_total_biom_3way)
```

# Root to shoot ratio

```{r}
lme_root_shoot_ratio_3way <- lme(root_shoot_ratio ~  treatment:nfixer:amax +
                                                        treatment:nfixer:gs  +
                                                        treatment:nfixer:wue +
                                                        treatment:nfixer:pnue +
                                                        #treatment:nfixer:d13c +

                                                        # Added
                                                        #treatment:nfixer:Narea_g_m2 +
                                                        init_height,
                            random = ~1 | spcode,
                            data = data_for_models_scaled)
```


## Model Selection

```{r}
dredge(lme_root_shoot_ratio_3way)
```


## R2

```{r}
r.squaredGLMM(lme_root_shoot_ratio_3way)
```

# Bootstraps

## Total Biomass

```{r}
if (file.exists("./processed_data/bootstrap_ci_total_biomass.RData")) {

    print("The bootstrap exists")

    bootstrap_ci_total_biomass <- readRDS("./processed_data/bootstrap_ci_total_biomass.RData")

    print("File read")

} else {

    print("The bootstrap does not exist, Creating bootstrap")

    bootstrap_ci_total_biomass <-
        bootstrap_model_ci_df(model = lme_total_biom_3way,
                              category = "all",

                              iter = 9999)

    saveRDS(bootstrap_ci_total_biomass,
            file = "./processed_data/bootstrap_ci_total_biomass.RData")

}
```


## Plot Total Biomass

```{r}

colmn <- paste0("col_", 1:3)

data_total_biomass_plot <-

        bootstrap_ci_total_biomass %>%

            unite("treat_trait", col_1, col_3, sep = "-", remove = F) %>%
            mutate(across(where(is.character), as.factor)) %>%
            filter(type == "norm") %>%
            mutate(treat_trait = factor(treat_trait, levels = c("plus_water_nutrients-amax",
                                                            #"plus_water_nutrients-d13c" ,
                                                            "plus_water_nutrients-gs"   ,
                                                            "plus_water_nutrients-pnue" ,
                                                            "plus_water_nutrients-wue",
                                                            "plus_water-amax" ,
                                                            #"plus_water-d13c",
                                                            "plus_water-gs"  ,
                                                            "plus_water-pnue",
                                                            "plus_water-wue",
                                                            "plus_nutrients-amax",
                                                            #"plus_nutrients-d13c",
                                                            "plus_nutrients-gs"  ,
                                                            "plus_nutrients-pnue",
                                                            "plus_nutrients-wue" ,
                                                            "no_additions-amax",
                                                            #"no_additions-d13c",
                                                            "no_additions-gs"  ,
                                                            "no_additions-pnue",
                                                            "no_additions-wue")))
```


```{r total_biomass_plot, fig.width = 16, fig.height = 13, fig.align = 'center'}

cleveland_plot(x = treat_trait, y = estimate,
                                     color = col_1,
                                     shape = significance,
                                     ci_lower = lower, ci_upper = upper,
                                     data = data_total_biomass_plot) +
        facet_wrap(. ~col_2)

```

