---
title: 'Model inference'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)

options(knitr.graphics.auto_pdf = TRUE, scipen = 9999, digits = 10)
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      options(digits = 3), 
                      # I added this option for edit the pdf in inkscape
                      dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,
                      
                      error          = FALSE,
                      fig.align      = "center",
                      
                      # Path where figures are going to be store pdf single figures
                      fig.path       = paste0("../model_inference/figures_model_inference", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# save a html copy file in a specific place
#doc.files <- c(list.files(pattern = "pdf"),
#               list.files(pattern = "html"),
#               list.files(pattern = "docx"))

#for (file in doc.files) {
#  cambiar nombre
#    file.rename(file, file.path("../../hw1/", file))
#}
```


```{r libaries, message=FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For tables appearance
library(reactablefmtr)
# For cleaning colnames
library(janitor)
# For add column function
library(tibble)
library(purrr)
library(performance)
# For bootstrap confidence intervals
library(lmeresampler)
library(nlme)
library(lme4)
# For model comparison
library(MuMIn)
# For get preditions
library(modelr)
# For reporting results
library(report)
library(emmeans)
```


```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
sys.source("./scripts/code_join_data_full_dataset.R", envir = knitr::knit_global())
```

```{r message=FALSE, warning=FALSE}
# Load functions

## Inference
sys.source("./R/functions_inference_anova_and_tukey_tables.R", envir = knitr::knit_global())

## Plots
sys.source("./R/function_masomenos_plots.R", envir = knitr::knit_global())
```

```{r}
# Load models
models_q1_q2_and_mass_fractions <- readRDS("./processed_data/models_q1_q2_and_mass_fractions.RData")

models_nodule_count <- readRDS("./processed_data/models_nodule_count.RData")

models_3way_q3 <- readRDS("./processed_data/models_3way_q3.RData")

```

# Models selected according validation plots 

## Models: Q1 and Q2

```{r}
names(models_q1_q2_and_mass_fractions)
```

## Nodule count seleted model

```{r}
names(models_nodule_count)
```

```{r echo = FALSE}

# Step done for getting predictions from model
source("./scripts/code_clean_data_nodules.R")

data_pred_nodules <-
     data_nodules_cleaned %>%
         
         # add id to rownames for keep track of the rows
         
         dplyr::select(c("id","spcode","treatment","init_height"))
```

## 3-way models

```{r}
names(models_3way_q3)
```

# Load data

```{r}
set.seed(666)
```

```{r echo=FALSE}
# Step done for getting predictions from models for Q1 and Q2 
data_for_models <- 
    data_for_models %>%
    
        rownames_to_column("id") %>% 
        
        # Remove unused variables 
        dplyr::select(id, spcode, treatment, nfixer, init_height)  

```

# Q1: Growth and biomass accumulation


## Anova and Tuckey comparisons 

```{r}
models_q1_q2_and_mass_fractions %>%
    
    # Remove traits and mass fractions models
    purrr::list_modify("rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       "d13c" = NULL, "amax" = NULL, "gs" = NULL, 
                       "wue_log" = NULL, "pnue_log" = NULL) %>% 
    anova_table_tidy(., model_list = T)
```

```{r}
models_q1_q2_and_mass_fractions %>%
    
    # Remove traits and mass fractions models
    purrr::list_modify("rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       "d13c" = NULL, "amax" = NULL, "gs" = NULL, 
                       "wue_log" = NULL, "pnue_log" = NULL) %>% 
    
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```

## Percentange change

```{r message=FALSE, warning=FALSE}
models_q1_q2_and_mass_fractions %>%
    
    # Remove traits and mass fractions models
    purrr::list_modify("rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       "d13c" = NULL, "amax" = NULL, "gs" = NULL, 
                       "wue_log" = NULL, "pnue_log" = NULL) %>% 
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```

## Plot Biomass growth boxplot

```{r message=FALSE, warning=FALSE}

# Get predictions
string <- c("models_q1_q2_and_mass_fractions")

data_pred_biomass_growth <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions of:
                            models_q1_q2_and_mass_fractions$rgr,
                            models_q1_q2_and_mass_fractions$total_biomass,
                            models_q1_q2_and_mass_fractions$above_biomass,
                            models_q1_q2_and_mass_fractions$below_biomass) %>%
            
            # Get fitted values
            pivot_wider(names_from = model, values_from = pred) %>% 
             rename_all(
              funs(
                  
                # rename columns
                stringr::str_to_lower(.) %>%
                
                # Remove string from name and replace it with pred_
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .))) %>%
        # Select only pred variables
        dplyr::select(-init_height)  
```

```{r warning=FALSE}

# Generate plot combinations
vars_q1 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_biomass_growth, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_biomass_growth, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_biomass_growth, treatment) %>%  names)
```

```{r fig_1_biomass_growth_boxplot, fig.height = 6, fig.width = 7, fig.align = 'center'}
vars_q1 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_biomass_growth,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) 
      #cowplot::plot_grid(plotlist = ., ncol = 2)
```

## r2 models

```{r}
models_q1_q2_and_mass_fractions %>%
    
    # Remove traits and mass fractions models
    purrr::list_modify("rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       
                       "d13c" = NULL, "amax" = NULL, "gs" = NULL, 
                       "wue_log" = NULL, "pnue_log" = NULL)%>%
    
    map(., r2) %>% 
    unlist() 
   
   
```

# Q2: Traits and Nodule count

## Anova and Tuckey comparisons 

```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL) %>%
    
    anova_table_tidy(., model_list = T)
```


### Models with significant interaction

```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       
                        "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       
                       # No interaction
                       "d13c" = NULL,
                       
                       # This only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```

### Models with signifcant treatment effect

```{r message=FALSE, warning=FALSE}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                        "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                        
                       "amax" = NULL,  "wue_log" = NULL,
                       "d13c" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment")
```
### Models with signifcant nfixer effect

```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                     
                       "amax" = NULL, "d15n" = NULL, 
                       "wue_log" = NULL, "d13c" = NULL, "pnue_log" = NULL) %>%
    
    tukey_table_tidy(., model_list = T, formula = "nfixer")
```

## Percentange change

### Models with significant interaction

```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       
                       
                       # No interaction
                       "d13c" = NULL, 
                       
                       # These variables only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```

### Models with significant treatment effect

```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                     
                       "amax" = NULL, "wue_log" = NULL
                       , "d13c" = NULL ) %>%
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment")
    

```


### Models with signifcant nfixer effect

```{r}

# Difference in gs between fixer non fixers
# Percentage difference
as.data.frame(emmeans(models_q1_q2_and_mass_fractions$gs,
                          specs = pairwise ~ "nfixer",
                          type = "response",
                          adjust = "tukey")$emmeans)  %>% 
        
        clean_names() %>%  
        dplyr::select(-c(df,lower_cl, upper_cl,se)) %>% 
        
        # Rename response to emmean, this is done when models is log  
        
       
        mutate(difference = ((emmean - first(emmean))),
               perc_difference =((emmean - first(emmean) )/first(emmean))*100) %>% 
        mutate_if(is.numeric, round, 3)
```


## Plot Traits boxplot

```{r}

string <- c("models_q1_q2_and_mass_fractions")

data_pred_traits <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions
                            models_q1_q2_and_mass_fractions$amax,
                            models_q1_q2_and_mass_fractions$d13c,
                            models_q1_q2_and_mass_fractions$gs,
                            models_q1_q2_and_mass_fractions$wue_log,
                            models_q1_q2_and_mass_fractions$pnue_log) %>% 
                       
            
            
        pivot_wider(names_from = model, values_from = pred) %>%
            rename_all(funs(
                
                # rename columns
                stringr::str_to_lower(.) %>%
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .)
                )) %>% 
    
        # Back transform log variables
        mutate(pred_wue = exp(pred_wue_log),
               pred_pnue = exp(pred_pnue_log)) %>%  
       
        # Remove log predictions and init height
        dplyr::select(-c(init_height, pred_wue_log,pred_pnue_log ))  
```


```{r}
vars_q2 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_traits, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_traits, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_traits, treatment) %>%  names)
```

```{r fig_2_traits_q2_boxplot, fig.height = 6, fig.width = 7, fig.align = 'center'}
vars_q2 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_traits,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) 
#      cowplot::plot_grid(plotlist = ., ncol = 2)
    
```



```{r}
models_q1_q2_and_mass_fractions %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL,
                       
                        "rmf_log" = NULL, "smf_log" = NULL, "lmf_log" = NULL,
                       ) %>%
    map(., r2) %>% 
    unlist() 
    
    
```


## Nodule count

```{r}
map(models_nodule_count, AIC)
```

```{r}
anova.lme(models_nodule_count$nlme_gaussian_log_weights, type = "marginal")
```




```{r}
# Interpretation: https://stats.oarc.ucla.edu/other/mult-pkg/faq/general/faqhow-do-i-interpret-a-regression-model-when-some-variables-are-log-transformed/
```


```{r}
as.data.frame(emmeans(models_nodule_count$nlme_gaussian_log_weights,
                           specs = pairwise ~ "treatment",
                           type = "response",
                           adjust =  "tukey")$contrast) %>%
         clean_names() %>% 
         reactable()
```


```{r}
# Percentage difference
emmeans_table_tidy(models_nodule_count$nlme_gaussian_log_weights, 
                        single_model = T, 
                        formula = "treatment")

```




```{r}
models_nodule_count %>% 
    #purrr::list_modify("rgr_slope" = NULL, "agr_log" = NULL, "gs_sqrt" = NULL) %>%
    map(., r2) %>% 
    unlist() 
```


# Q3: Do species modulate their traits in response to the addition of nutrients and/or water to increase biomass accumulation and growth rates? 

For answering this question I fitted the following models focusing in the 3 way interaction:

$$response \sim\ treatment:nfixer:trait\ + initialheight\ + (1|spcode) $$

## r2

```{r}
map(models_3way_q3, r.squaredGLMM)
```

## Bootstraps

### Total Biomass
 
```{r}
# Look results before bootstrap
#report::report(models_3way_q3$lme_total_biom_3way)
```


```{r}
if (file.exists("./processed_data/bootstrap_ci_total_biomass.RData")) {
    
    print("The bootstrap exists")
    
    bootstrap_ci_total_biomass <- readRDS("./processed_data/bootstrap_ci_total_biomass.RData")
    
    print("File read")
    
} else {
    
    print("The bootstrap does not exist, Creating bootstrap")
    
    bootstrap_ci_total_biomass <- 
        bootstrap_model_ci_df(model = models_3way_q3$lme_total_biom_3way, 
                              category = "all", 
                              iter = 9999) 
    
    saveRDS(bootstrap_ci_total_biomass, 
            file = "./processed_data/bootstrap_ci_total_biomass.RData")
    
}
```


#### Plot Total Biomass

```{r}
colmn <- paste0("col_", 1:3)

data_total_biomass_plot <- 
        
        bootstrap_ci_total_biomass %>%
            tidyr::separate(data = ., col = term, 
                     sep = ":",
                     into = colmn,
                     remove = F) %>% 
            mutate(across(where(is.character), as.factor)) %>% 
            filter(type == "norm")   

data_total_biomass_plot$term <- 
    factor(data_total_biomass_plot$term,
           levels = data_total_biomass_plot$term[order(data_total_biomass_plot$col_2)])

# Show estimates
data_total_biomass_plot %>% 
    arrange(desc(significance)) %>% 
    dplyr::select(term,estimate,lower, upper, response_var,significance)
```


```{r total_biomass_plot, fig.width = 30, fig.height = 25, fig.align = 'center'}
# Plot
# 

(total_biomass_plot <- cleveland_plot(x = term, y = estimate, shape = col_2,
                                     color = significance,
                                     ci_lower = lower, ci_upper = upper,
                                     data = data_total_biomass_plot))

 
```

### Belowground Biomass

```{r}
# Look results before bootstrap
#report::report_parameters(models_3way_q3$lme_below_biom_3way)
```



```{r}
if (file.exists("./processed_data/bootstrap_ci_belowground_biomass.RData")) {
    
    print("The bootstrap exists")
    
    bootstrap_ci_belowground_biomass <- readRDS("./processed_data/bootstrap_ci_belowground_biomass.RData")
    
    print("File read")
    
} else {
    
    print("The bootstrap does not exist, Creating bootstrap")
    
    bootstrap_ci_belowground_biomass <- 
        bootstrap_model_ci_df(model = models_3way_q3$lme_below_biom_3way, 
                              category = "all", 
                              iter = 9999) 
       
    saveRDS(bootstrap_ci_belowground_biomass, 
            file = "./processed_data/bootstrap_ci_belowground_biomass.RData")
}
```

#### Plot Belowground Biomass

```{r}

colmn <- paste0("col_", 1:3)

data_belowground_plot <-
    
    # Clean data
        bootstrap_ci_belowground_biomass %>% 
            tidyr::separate(data = ., col = term, 
                     sep = ":",
                     into = colmn,
                     remove = F) %>% 
            mutate(across(where(is.character), as.factor)) %>% 
            filter(type == "norm") 

# Rearrange data
data_belowground_plot$term <- 
    factor(data_belowground_plot$term,
           levels = data_belowground_plot$term[order(data_belowground_plot$col_2)])



# Show estimates
data_belowground_plot %>% 
    arrange(desc(significance)) %>% 
    dplyr::select(term,estimate,lower, upper, response_var,significance)


```


```{r belowground_plot, fig.width = 30, fig.height = 25, fig.align = 'center', echo = FALSE}
# Plot
belowground_plot <- cleveland_plot(x = term, y = estimate, shape = col_2,
                                   color = significance,
                                   ci_lower = lower, ci_upper = upper,
                                   data = data_belowground_plot)

 
```




### Aboveground Biomass

```{r}
# Look results before bootstrap
#report::report_parameters(models_3way_q3$lme_above_biom_3way)
```

```{r}
# Look results before bootstrap
AIC(models_3way_q3$lme_above_biom_3way)

data.frame(intervals(models_3way_q3$lme_above_biom_3way)$fixed) %>% 
    tibble::rownames_to_column("names") %>% 
    filter(!names %in% c("(Intercept)","init_height")) %>% 
    mutate(significance = ifelse((lower > 0 & upper > 0 | lower < 0 & upper < 0 ), 
                                 TRUE, FALSE)) %>% 
    filter(significance == TRUE) %>% 
    reactable()
```

```{r}
if (file.exists("./processed_data/bootstrap_ci_aboveground_biomass.RData")) {
    
    print("The bootstrap exists")
    
    bootstrap_ci_aboveground_biomass <- readRDS("./processed_data/bootstrap_ci_aboveground_biomass.RData")
    
    print("File read")
    
} else {
    
    print("The bootstrap does not exist, Creating bootstrap")
    
    bootstrap_ci_aboveground_biomass <- 
        bootstrap_model_ci_df(model = models_3way_q3$lme_above_biom_3way, 
                              category = "all", 
                              iter = 9999)
        
    saveRDS(bootstrap_ci_aboveground_biomass, 
            file = "./processed_data/bootstrap_ci_aboveground_biomass.RData")
}
```
#### Plot Aboveground Biomass

```{r}
colmn <- paste0("col_", 1:3)

data_aboveground_plot <- 
        
        # Clean data
        bootstrap_ci_aboveground_biomass %>%
                tidyr::separate(data = ., col = term, 
                                                sep = ":",
                                                into = colmn,
                                            remove = F) %>% 
                mutate(across(where(is.character), as.factor))  %>% 
                filter(type == "norm") 

# Rearrange data
data_aboveground_plot$term <- 
    factor(data_aboveground_plot$term,
           levels = data_aboveground_plot$term[order(data_aboveground_plot$col_2)])

data_aboveground_plot %>% 
    arrange(desc(significance)) %>% 
    dplyr::select(term,estimate,lower, upper, response_var,significance)
```


```{r aboveground_plot, fig.width = 30, fig.height = 25, fig.align = 'center', echo = FALSE}
# Plot
(aboveground_plot <- cleveland_plot(x = term, y = estimate, shape = col_2, color = significance,
               ci_lower = lower, ci_upper = upper,
               data = data_aboveground_plot))

```



### RGR

```{r}
# Look results before bootstrap
#report::report_parameters(models_3way_q3$lme_rgr_3way)
```


```{r}
# Look results before bootstrap

AIC(models_3way_q3$lme_rgr_3way)

data.frame(intervals(models_3way_q3$lme_rgr_3way)$fixed) %>% 
    tibble::rownames_to_column("names") %>% 
    filter(!names %in% c("(Intercept)","init_height")) %>% 
    mutate(significance = ifelse((lower > 0 & upper > 0 | lower < 0 & upper < 0 ), 
                                 TRUE, FALSE)) %>% 
    filter(significance == TRUE)  %>% 
    reactable()
```


```{r}
if (file.exists("./processed_data/bootstrap_ci_rgr.RData")) {
    
    print("The bootstrap exists")
    
    bootstrap_ci_rgr <- readRDS("./processed_data/bootstrap_ci_rgr.RData")
    
    print("File read")
    
} else {
    
    print("The bootstrap does not exist, Creating bootstrap")

    bootstrap_ci_rgr <- 
        bootstrap_model_ci_df(model = models_3way_q3$lme_rgr_3way, 
                              category = "all", 
                              iter = 9999) 
    saveRDS(bootstrap_ci_rgr, 
            file = "./processed_data/bootstrap_ci_rgr.RData")
    
}
```

#### Plot RGR
```{r}
colmn <- paste0("col_", 1:3)

data_rgr_plot <- 
    bootstrap_ci_rgr %>% 
            tidyr::separate(data = ., col = term, 
                     sep = ":",
                     into = colmn,
                     remove = F) %>% 
            mutate(across(where(is.character), as.factor)) %>% 
            filter(type == "norm")

data_rgr_plot$term <- 
    factor(data_rgr_plot$term,
           levels = data_rgr_plot$term[order(data_rgr_plot$col_2)])
           
        
# Plot
rgr_plot <- cleveland_plot(x = term, y = estimate, shape = col_2,
                           color = significance,
                           ci_lower = lower, ci_upper = upper,
                             data = data_rgr_plot)

```

 
##
```{r figure_3_three_way_q3, fig.width = 40, fig.height = 30, fig.align = 'center', echo = FALSE}
cowplot::plot_grid(total_biomass_plot,belowground_plot,aboveground_plot,rgr_plot, ncol = 2)
```
 
 