---
title: 'Model inference'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)

options(knitr.graphics.auto_pdf = TRUE, scipen = 9999, digits = 10)
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      options(digits = 3), 
                      # I added this option for edit the pdf in inkscape
                      dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,
                      
                      error          = FALSE,
                      fig.align      = "center",
                      
                      # Path where figures are going to be store pdf single figures
                      fig.path       = paste0("../model_inference/figures_model_inference", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# save a html copy file in a specific place
#doc.files <- c(list.files(pattern = "pdf"),
#               list.files(pattern = "html"),
#               list.files(pattern = "docx"))

#for (file in doc.files) {
#  cambiar nombre
#    file.rename(file, file.path("../../hw1/", file))
#}
```


```{r libaries, message=FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For tables appearance
library(reactablefmtr)
# For cleaning colnames
library(janitor)
# For add column function
library(tibble)
library(purrr)
library(performance)
# For bootstrap confidence intervals
library(lmeresampler)
library(nlme)
library(lme4)
# For model comparison
library(MuMIn)
# For get preditions
library(modelr)
# For reporting results
library(report)
library(emmeans)
```


```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
sys.source("./scripts/code_join_data_full_dataset.R", envir = knitr::knit_global())
```

```{r message=FALSE, warning=FALSE}
# Load functions

## Inference
sys.source("./R/functions_inference_anova_and_tukey_tables.R", envir = knitr::knit_global())

## Plots
sys.source("./R/function_masomenos_plots.R", envir = knitr::knit_global())
```

```{r}
# Load models
models_q1_q2 <- readRDS("./processed_data/models_q1_q2.RData")

models_q3_lmer <- readRDS("./processed_data/models_q3_3_way_interaction_lmer.RData")

models_q3_nlme <- readRDS("./processed_data/models_q3_3_way_interaction_nlme.RData")

models_nodule_count <- readRDS("./processed_data/models_nodule_count.RData")
```

# Models selected according validation plots for answering questions

## Models Q1 and Q2

```{r}
models_q1_q2 <-     
    models_q1_q2 %>%
        
        # Remove traits models not log transformed
        purrr::list_modify("wue" = NULL, "pnue"  = NULL)

names(models_q1_q2)

```

## Nodule count seleted model

```{r}
models_nodule_count <- 
    models_nodule_count %>%
        
        # Remove unused models
        purrr::list_modify("lmer_gaussian_log" = NULL, "glmer_poisson"  = NULL)
```



# Double checking nodule count model's assumptions
```{r}
plot(models_nodule_count$lmer_gaussian, resid(., scaled=TRUE) ~ fitted(.), 
     abline = 0,pch=16,xlab="Fitted values",ylab="Standardised residuals")
```

```{r echo = FALSE}
source("./scripts/code_clean_data_nodules.R")
# Delete unused variables
data_nodules_cleaned <-
    data_nodules_cleaned %>%
        
        # add id to rownames for keep track of the rows
        column_to_rownames("id") %>% 
        dplyr::select(spcode,treatment, everything())
```


```{r}
plot(models_nodule_count$lmer_gaussian, resid(., scaled=TRUE) ~ fitted(.), 
     abline = 0,pch = 16, col = data_nodules_cleaned$spcode, 
     xlab = "Fitted values", ylab = "Standardised residuals")
```

```{r}
plot(models_nodule_count$lmer_gaussian, 
     resid(., scaled=TRUE) ~ fitted(.)| spcode, abline = 0, pch=16, 
     xlab = "Fitted values",ylab = "Standardised residuals")
```

```{r}
plot(models_nodule_count$lmer_gaussian, spcode ~ resid(., scaled=TRUE),abline=0, 
     pch=16, xlab="Standardised residuals", ylab="Regions")
```

```{r}
lev <- hat(model.matrix(models_nodule_count$lmer_gaussian))
plot(resid(models_nodule_count$lmer_gaussian,type = "pearson")~lev,
     las = 1, ylab = "Standardised residuals", xlab = "Leverage")

```
 
```{r}
# Calculate Cook's Distance
# N.B. If Cook's distance is greater than 1 this highlights problematic datapoints
cd <- cooks.distance(models_nodule_count$lmer_gaussian)

# Plot leverage and Cook's distance together
par(mfrow = c(1,1))

plot(lev, pch = 16, col = "red", ylim = c(0,1.2), las = 1, 
     ylab = "Leverage/Cook's distance value")

points(cd, pch = 17, col = "blue")

points(x = 150, y = 1.1, pch = 16, col = "red")

points(x = 150, y = 0.9,pch = 17,col = "blue")

text(x = 155,y = 1.1, "Leverage", adj = c(0,0.5))

text(x = 155, y = 0.9, "Cook's distance", adj = c(0,0.5))
```


# Load data

```{r}
set.seed(666)
```

```{r echo=FALSE}
# Step done for getting predictions from models for Q1 and Q2 
data_for_models <- 
    data_for_models %>% 
        
        # Remove unused variables 
        dplyr::select(spcode, treatment, nfixer, init_height, everything(),
               -c(8,10))  

```

# Q1 Growth and biomass accumulation

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    anova_table_tidy(., model_list = T)
```

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```


```{r message=FALSE, warning=FALSE}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```

## Plot Biomass growth boxplot

```{r message=FALSE, warning=FALSE}

# Get predictions
string <- c("models_q1_q2")

data_pred_biomass_growth <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions of:
                            models_q1_q2$rgr,
                            models_q1_q2$total_biomass,
                            models_q1_q2$above_biomass,
                            models_q1_q2$below_biomass) %>%
            
            # Get fitted values
            pivot_wider(names_from = model, values_from = pred) %>% 
             rename_all(
              funs(
                  
                # rename columns
                stringr::str_to_lower(.) %>%
                
                # Remove string from name and replace it with pred_
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .))) %>% 
    
        # Select only pred variables
        dplyr::select(c(1:3,15:18))  
```

```{r warning=FALSE}

# Generate plot combinations
vars_q1 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_biomass_growth, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_biomass_growth, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_biomass_growth, treatment) %>%  names)
```


```{r fig_1_biomass_growth_boxplot, fig.height = 6, fig.width = 7, fig.align = 'center'}
vars_q1 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_biomass_growth,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) 
      #cowplot::plot_grid(plotlist = ., ncol = 2)
```


## r2 models

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>%
    
    map(., r2)
```

# Q2 Traits and Nodule count

## Anova and Tuckey comparisons 

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL) %>%
    
    anova_table_tidy(., model_list = T)
```


### Models with significant interaction

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       # No interaction
                       "d13c" = NULL,
                       
                       # This only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```

### Models with no significant interaction

```{r message=FALSE, warning=FALSE}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                     
                       "amax" = NULL, "d15n" = NULL, "wue_log" = NULL
                       , "d13c" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment")
```


```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                     
                       "amax" = NULL, "d15n" = NULL, 
                       "wue_log" = NULL, "d13c" = NULL, "pnue_log" = NULL) %>%
    
    tukey_table_tidy(., model_list = T, formula = "nfixer")
```

## Percentange change

### Models with significant interaction

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       # No interaction
                       "d13c" = NULL, 
                       
                       # These variables only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```

### Models with no significant interaction

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                     
                       "amax" = NULL, "wue_log" = NULL
                       , "d13c" = NULL ) %>%
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment")
    

```
```{r}

# Difference in WUE btw fixer non fixers
# Percentage difference
as.data.frame(emmeans(models_q1_q2$gs,
                          specs = pairwise ~ "nfixer",
                          type = "response",
                          adjust = "tukey")$emmeans)  %>% 
        
        clean_names() %>%  
        dplyr::select(-c(df,lower_cl, upper_cl,se)) %>% 
        
        # Rename response to emmean, this is done when models is log  
        
       
        mutate(difference = ((emmean - first(emmean))),
               perc_difference =((emmean - first(emmean) )/first(emmean))*100) %>% 
        mutate_if(is.numeric, round, 3)
```


## Plot Traits boxplot

```{r}
string <- c("models_q1_q2")

data_pred_traits <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions
                            models_q1_q2$amax,
                            models_q1_q2$d13c,
                            models_q1_q2$gs,
                            models_q1_q2$wue_log,
                            models_q1_q2$pnue_log) %>% 
                       
            
            
        pivot_wider(names_from = model, values_from = pred) %>%
            rename_all(funs(
                
                # rename columns
                stringr::str_to_lower(.) %>%
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .)
                )) %>%
    
        # Back transform log variables
        mutate(pred_wue = exp(pred_wue_log),
               pred_pnue = exp(pred_pnue_log)) %>% 
       
        dplyr::select(c(1:3,15:17, 20,21))  
```


```{r}
vars_q2 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_traits, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_traits, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_traits, treatment) %>%  names)
```


```{r fig_2_traits_q2_boxplot, fig.height = 6, fig.width = 7, fig.align = 'center'}
vars_q2 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_traits,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) 
#      cowplot::plot_grid(plotlist = ., ncol = 2)
    
```

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL) %>%
    
    map(., r2)
```


## Nodule count

```{r}
anova_table_tidy(models_nodule_count$lmer_gaussian, single_model = T)
```


```{r}
arm::display(models_nodule_count$lmer_gaussian)
sjPlot::plot_model(models_nodule_count$lmer_gaussian)
```

```{r}
tukey_table_tidy(models_nodule_count$lmer_gaussian,single_model = T, 
                 formula = "treatment")
```


```{r}
models_nodule_count %>% 
    #purrr::list_modify("rgr_slope" = NULL, "agr_log" = NULL, "gs_sqrt" = NULL) %>%
    map(., r2)
```


# Q3 Do species modulate their traits in response to the addition of nutrients and/or water to increase biomass accumulation and growth rates? 

For answering this question I fitted the following models focusing in the 3 way interaction:

$$response \sim\ treatment:nfixer:trait\ + initialheight\ + (1|spcode) $$
I have 4 response variables(rgr, total biomass, above and below ground biomass) and 6 traits thus I fitted 24 models. Also I scaled (z-scores) the traits for being able to compare between parameters of different models.  

After this I bootstrap the parameters for getting the Confidence intervals

## Model comparisons

### Aboveground biomass

```{r}
model.sel(models_q3_nlme$`abovebiomass~amax`,
          models_q3_nlme$`abovebiomass~d13c`,
          models_q3_nlme$`abovebiomass~gs`,
          models_q3_nlme$`abovebiomass~pnue`,
          models_q3_nlme$`abovebiomass~wue`)[,8:12]

```


```{r}
model.sel(models_q3_nlme$`abovebiomass~amax`,
          models_q3_nlme$`abovebiomass~d13c`,
          models_q3_nlme$`abovebiomass~gs`,
          models_q3_nlme$`abovebiomass~pnue`,
          models_q3_nlme$`abovebiomass~wue`,
          rank = "AIC")[,8:12]
```


### Belowgraound Biomass

```{r}
model.sel(models_q3_nlme$`belowbiomass~amax`,
                    #models_q3_nlme$`belowbiomass~d15n`,
                    models_q3_nlme$`belowbiomass~d13c`,
                    models_q3_nlme$`belowbiomass~gs`,
                    models_q3_nlme$`belowbiomass~pnue`,
                    models_q3_nlme$`belowbiomass~wue`,
                    rank = "AIC")[,8:12]
```

### rgr

```{r}
model.sel(models_q3_nlme$`rgr~amax`,
                    models_q3_nlme$`rgr~d13c`,
                    #models_q3_nlme$`rgr~d15n`,
                    models_q3_nlme$`rgr~gs`,
                    models_q3_nlme$`rgr~pnue`,
                    models_q3_nlme$`rgr~wue`)[,8:12]
```


### Total Biomass

```{r}
model.sel(models_q3_nlme$`totalbiomass~amax`,
                   
                    models_q3_nlme$`totalbiomass~d13c`,
                    models_q3_nlme$`totalbiomass~gs`,
                    models_q3_nlme$`totalbiomass~pnue`,
                    models_q3_nlme$`totalbiomass~wue`)[,8:12]
```

## r2

```{r}
map(models_q3_nlme, r.squaredGLMM)
```

## Anova tables

```{r}
models_q3_nlme %>%
    anova_table_tidy(., model_list = T)
```

## Bootstrap parameters for getting confidence intervals

```{r bootstrap__ci_three_way_params_models}

# Test bootstrap_model_ci_df function
# bootstrap_model_ci_df(model = models_q3_nlme$`abovebiomass~amax`,
#                       category = "all", iter = 100)

# bootstrap_ci_results <- map_dfr(models_q3_nlme,~ bootstrap_model_ci_df(model = .x,
#                                                                category = "all",
#                                                                iter = 9999))
# 
# # Save
# saveRDS(bootstrap_ci_results, file = "./processed_data/bootstrap_ci_results.RData")

# Read dataset
bootstrap_ci_results <- readRDS("./processed_data/bootstrap_ci_results.RData")
```




## Plot

### Only significant coefficients

```{r figure_3_three_way_q3, fig.width = 32, fig.height = 20, fig.align = 'center', echo = FALSE}

colmn <- paste0("col_", 1:3)
bootstrap_ci_results %>% 
    tidyr::separate(data = ., col = term, 
                    sep = ":",
                    into = colmn,
                    remove = F) %>% 
    mutate(across(where(is.character), as.factor)) %>% 
    filter(type == "norm" & significance == TRUE)  %>% 

    ggplot2::ggplot(data = ., aes(x = term, y = estimate, color = significance,
                                  shape = col_2)) +
            ggplot2::geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) +
            ggplot2::geom_point(size = 7, position = position_dodge(width = .9)) +

            # 95% C.I
            ggplot2::geom_linerange(aes(ymin = lower, ymax = upper),
                                    lwd = 1, position = position_dodge(width = .9)) +

            ggplot2::theme_bw() +

            ggplot2::theme(legend.position = "none",
                           strip.text.x = element_text(size = 25),
                           axis.text.y   = element_text(size= 25),
                           axis.text.x   = element_text(size= 25),
                           axis.title.y  = element_text(size= 25),
                           axis.title.x  = element_text(size= 25),
                           panel.grid.major.y = element_blank(),
                           panel.grid.minor = element_blank(),
                           axis.line = element_line(size = .4,colour = "black"),
                           panel.border = element_rect(colour = "black", fill= NA,
                                                       size = 1.3)) +

            # Significance colors
            ggplot2::scale_colour_manual(values = c("black")) +
            ggplot2::ylab("Estimated coefficients (median +/- 95CI)") +
            ggplot2::xlab("") +

            ggplot2::facet_wrap( ~ response_var, scales = "free", ncol = 2) +

            ggplot2::coord_flip()

```


### All coefficients

```{r supplementary_figure_3_three_way_q3, fig.width = 30, fig.height = 25, fig.align = 'center', echo = FALSE}
colmn <- paste0("col_", 1:3)
bootstrap_ci_results %>% 
    tidyr::separate(data = ., col = term, 
                    sep = ":",
                    into = colmn,
                    remove = F) %>% 
    mutate(across(where(is.character), as.factor)) %>% 
    filter(type == "norm")  %>% 


ggplot2::ggplot(data = .,
               # Highlight significant terms
               aes(x = term, y = estimate, color = significance,  shape = col_2)) +


            ggplot2::geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) +
            ggplot2::geom_point(size = 6, position = position_dodge(width = .9)) +

            # 95% C.I
            ggplot2::geom_linerange(aes(ymin = lower, ymax = upper),
                                    lwd = 1, position = position_dodge(width = .9)) +

            # 99% C.I
            #geom_linerange(aes(ymin = !!yvar - !!se * interval2,
            #                   ymax = !!yvar + !!se * interval2),
            #               lwd = 1/2, position = position_dodge(width = .9)) +

            ggplot2::theme_bw() +

            ggplot2::theme(legend.position = "none",
                           strip.text.x = element_text(size = 25),
                           axis.text.y   = element_text(size= 25),
                           axis.text.x   = element_text(size= 25),
                           axis.title.y  = element_text(size= 25),
                           axis.title.x  = element_text(size= 25),
                           panel.grid.major.y = element_blank(),
                           panel.grid.minor = element_blank(),
                           axis.line = element_line(size = .4,colour = "black"),
                           panel.border = element_rect(colour = "black", fill= NA,
                                                       size = 1.3)) +
            # Add name
            #labs(title = paste0(response_variable)) +

            # Significance colors
            ggplot2::scale_colour_manual(values = c("grey","black")) +
            ggplot2::ylab("Estimated coefficients (median +/- 95CI)") +
            ggplot2::xlab("") +

            ggplot2::facet_wrap(~ response_var, scales = "free", ncol = 2) +


            ggplot2::coord_flip()

```
