---
title: 'Model inference'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)

options(knitr.graphics.auto_pdf = TRUE, scipen = 9999)
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      options(digits = 3), 
                      # I added this option for edit the pdf in inkscape
                      dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,
                      
                      error          = FALSE,
                      fig.align      = "center",
                      
                      # Path where figures are going to be store pdf single figures
                      fig.path       = paste0("../model_inference/figures_model_inference_lmer", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# save a html copy file in a specific place
#doc.files <- c(list.files(pattern = "pdf"),
#               list.files(pattern = "html"),
#               list.files(pattern = "docx"))

#for (file in doc.files) {
#  cambiar nombre
#    file.rename(file, file.path("../../hw1/", file))
#}
```


```{r libaries, message=FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For tables appearance
library(reactablefmtr)
# For cleaning colnames
library(janitor)
# For add column function
library(tibble)
library(purrr)
library(performance)
# For bootstrap confidence intervals
library(lmeresampler)
# For model comparison
library(MuMIn)
# For get preditions
library(modelr)
# For reporting results
library(report)
library(emmeans)
```


```{r message=FALSE, warning=FALSE}
# Load dataset for adding predictions
sys.source("./scripts/code_join_data_full_dataset.R", envir = knitr::knit_global())
```

```{r message=FALSE, warning=FALSE}
# Load functions

## Inference
sys.source("./R/functions_inference_anova_and_tukey_tables.R", envir = knitr::knit_global())

## Plots
sys.source("./R/function_masomenos_plots.R", envir = knitr::knit_global())
```

```{r}
# Load models
models_q1_q2 <- readRDS("./processed_data/models_q1_q2.RData")

models_q3 <- readRDS("./processed_data/models_q3_3_way_interaction.RData")

models_q3_nlme <- readRDS("./processed_data/models_q3_3_way_interaction_nlme.RData")

models_nodule_count <- readRDS("./processed_data/models_list_nodule_count.RData")
```

# Models selected according validation plots for answering questions

## Models Q1 and Q2

```{r}
models_q1_q2 <-     
    models_q1_q2 %>%
        
        # Remove traits models
        purrr::list_modify("wue" = NULL, "pnue"  = NULL)
```


```{r}
models_nodule_count <- 
    models_nodule_count %>%
        
        # Remove traits models
        purrr::list_modify("lmer_gaussian_log" = NULL, "glmer_poisson"  = NULL)
```


## Models Q3 

```{r}
models_q3_nlme <- models_q3_nlme
```


# Load data

```{r}
set.seed(666)
```

```{r echo=FALSE}
# Step done for getting predictions from models
data_for_models <- 
    data_for_models %>% 
        
        # Remove unused variables 
        dplyr::select(spcode, treatment, nfixer, init_height, everything(),
               -c(8,10))  

```

# Q1 Growth and biomass accumulation

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "d15n"  = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    anova_table_tidy(., model_list = T)
```

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL, "d15n"  = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```


```{r message=FALSE, warning=FALSE}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL,"d15n"  = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>% 
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```

## Plot Biomass growth boxplot

```{r message=FALSE, warning=FALSE}
string <- c("models_q1_q2")

data_pred_biomass_growth <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions of:
                            models_q1_q2$rgr,
                            models_q1_q2$total_biomass,
                            models_q1_q2$above_biomass,
                            models_q1_q2$below_biomass) %>%
            
            # Get fitted values
            pivot_wider(names_from = model, values_from = pred) %>% 
             rename_all(
              funs(
                  
                # rename columns
                stringr::str_to_lower(.) %>%
                
                # Remove string from name and replace it with pred_
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .))) %>% 
    
        # Select only pred variables
        dplyr::select(c(1:3,15:18))  
```

```{r warning=FALSE}
vars_q1 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_biomass_growth, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_biomass_growth, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_biomass_growth, treatment) %>%  names)
```


```{r fig_1_biomass_growth_boxplot, fig.height = 7, fig.width = 7, fig.align = 'center'}
vars_q1 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_biomass_growth,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) %>% 
      cowplot::plot_grid(plotlist = ., ncol = 2)
```


## Models r2

```{r}
models_q1_q2 %>%
    
    # Remove traits models
    purrr::list_modify("d13c" = NULL,"d15n"  = NULL, "amax" = NULL, 
                       "gs" = NULL, "wue_log" = NULL, "pnue_log" = NULL) %>%
    
    map(., r2)
```

# Q2 Traits and Nodule count

## Anova and Tuckey comparisons 

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL) %>%
    
    anova_table_tidy(., model_list = T)
```

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       # No interaction
                       "d13c" = NULL,
                       
                       # This only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment|nfixer")
```

```{r message=FALSE, warning=FALSE}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                     
                       "amax" = NULL, "d15n" = NULL, "wue_log" = NULL
                       , "d13c" = NULL ) %>%
    
    tukey_table_tidy(., model_list = T, formula = "treatment")
```


```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                     
                       "amax" = NULL, "d15n" = NULL, 
                       "wue_log" = NULL, "d13c" = NULL, "pnue_log" = NULL) %>%
    
    tukey_table_tidy(., model_list = T, formula = "nfixer")
```

## Percentanfe change

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL, 
                       
                       # No interaction
                       "d13c" = NULL, 
                       
                       # This only responded to the treatments 
                       "pnue_log" = NULL, "gs" = NULL ) %>%
    
    # Percentage difference
    emmeans_table_tidy(., model_list = T, 
                        formula = "treatment|nfixer", 
                        grouping_var = "nfixer")
```


## Plot Traits boxplot

```{r}
string <- c("models_q1_q2")

data_pred_traits <- 
        
        # Get models prediction
        gather_predictions(data_for_models,
                           
                           # Return predictions
                            models_q1_q2$d15n,
                            models_q1_q2$amax,
                            models_q1_q2$d13c,
                            models_q1_q2$gs,
                            models_q1_q2$wue,
                            models_q1_q2$pnue) %>% 
                       
            
            
        pivot_wider(names_from = model, values_from = pred) %>%
            rename_all(funs(
                
                # rename columns
                stringr::str_to_lower(.) %>%
                stringr::str_replace(., c(string),"pred_") %>% 
                
                # Remove dollar sing    
                gsub("\\$", "", .)
                )) %>% 
       
        dplyr::select(c(1:3,15:20))  
```


```{r}
vars_q2 <-  
  crossing(
    
    # Get all numeric variables to plot (all y)
    as_tibble(t(combn(dplyr::select(data_pred_traits, where(is.numeric)) %>% names, 1))),
    
    # Select factor variables to plot
    x_axis_var = dplyr::select(data_pred_traits, nfixer) %>%  names,
    group_var = dplyr::select(data_pred_traits, treatment) %>%  names)
```


```{r fig_2_traits_q2_boxplot, fig.height = 14, fig.width = 12, fig.align = 'center'}
vars_q2 %>% 
      # Gererate plots
      pmap( ~ boxplot_plot_pmap(data = data_pred_traits,
                                y = !!sym(..1), x = !!sym(..2), 
                                fill = !!sym(..3))) %>% 
      cowplot::plot_grid(plotlist = ., ncol = 2)
    
```

```{r}
models_q1_q2 %>% 
    
    # Remove performance variables
    purrr::list_modify("total_biomass" = NULL, "above_biomass"  = NULL, 
                       "below_biomass" = NULL, "rgr" = NULL) %>%
    
    map(., r2)
```


## Nodule count

```{r}
anova_table_tidy(models_nodule_count$lmer_gaussian, single_model = T)
```
```{r}
arm::display(models_nodule_count$lmer_gaussian)
sjPlot::plot_model(models_nodule_count$lmer_gaussian)
```

```{r}
tukey_table_tidy(models_nodule_count$lmer_gaussian,single_model = T, 
                 formula = "treatment")
```


```{r}
models_nodule_count %>% 
    #purrr::list_modify("rgr_slope" = NULL, "agr_log" = NULL, "gs_sqrt" = NULL) %>%
    map(., r2)
```


# Q3 Do species modulate their traits in response to the addition of nutrients and/or water to increase biomass accumulation and growth rates? 

For answering this question I fitted the following models focusing in the 3 way interaction:

$$response \sim\ treatment:nfixer:trait\ + initialheight\ + (1|spcode) $$
I have 4 response variables(rgr, total biomass, above and below ground biomass) and 6 traits thus I fitted 24 models. Also I scaled (z-scores) the traits for being able to compare between parameters of different models.  

After this I bootstrap the parameters for getting the Confidence intervals

## Model comparisons

### Aboveground biomass

```{r}
model.sel(models_q3_nlme$`abovebiomass~amax`,
          models_q3_nlme$`abovebiomass~d15n`,
          models_q3_nlme$`abovebiomass~d13c`,
          models_q3_nlme$`abovebiomass~gs`,
          models_q3_nlme$`abovebiomass~pnue`,
          models_q3_nlme$`abovebiomass~wue`)[,9:13]
```


### Belowgraound Biomass

```{r}
model.sel(models_q3_nlme$`belowbiomass~amax`,
                    models_q3_nlme$`belowbiomass~d15n`,
                    models_q3_nlme$`belowbiomass~d13c`,
                    models_q3_nlme$`belowbiomass~gs`,
                    models_q3_nlme$`belowbiomass~pnue`,
                    models_q3_nlme$`belowbiomass~wue`)[,9:13]
```

### rgr

```{r}
model.sel(models_q3_nlme$`rgr~amax`,
                    models_q3_nlme$`rgr~d13c`,
                    models_q3_nlme$`rgr~d15n`,
                    models_q3_nlme$`rgr~gs`,
                    models_q3_nlme$`rgr~pnue`,
                    models_q3_nlme$`rgr~wue`)[,9:13]
```


### Total Biomass

```{r}

model.sel(models_q3_nlme$`totalbiomass~amax`,
                    models_q3_nlme$`totalbiomass~d15n`,
                    models_q3_nlme$`totalbiomass~d13c`,
                    models_q3_nlme$`totalbiomass~gs`,
                    models_q3_nlme$`totalbiomass~pnue`,
                    models_q3_nlme$`totalbiomass~wue`)[,9:13]
```

## r2

```{r}
map(models_q3_nlme, r.squaredGLMM)
```

## Anova tables

```{r}
#models_q3 %>%
#    anova_table_tidy(., model_list = T)
```

## Bootstrap parameters for getting confidence intervals

```{r}

a <- bootstrap_model_df(model = models_q3_nlme$`abovebiomass~amax`, 
                        category = "norm", iter = 100) 

```


```{r}
#.f = fixef
#

library(lmeresampler)

lmeresampler::bootstrap(models_q3$`abovebiomass~amax`,.f = fixef, B = 100, type = "parametric")
confint()
```


```{r}
confint(parametric_bootstrap(models_q3_nlme$`abovebiomass~amax`,.f = fixef, B = 100))
```

```{r}
library(lme4)
vcmodA <- lme4::lmer(mathAge11 ~ mathAge8 + gender + class + (1 | school), data = jsp728)
lmeresampler::bootstrap(vcmodA, .f = fixef, type = "parametric", B = 100)
bootstrap(model = vcmodA, .f = fixef,B = 100 )
```

```{r bootstrap_three_way_params, cache = TRUE}
#three_way_param_bootstrap <- map_dfr(models_q3, ~ bootstrap_model_df(., iter = 9999))
```

## Plot

```{r figure_3_three_way_q3, fig.width = 30, fig.height = 25, fig.align = 'center'}

colmn <- paste0("col_", 1:3)
three_way_param_bootstrap %>% 
    tidyr::separate(data = ., col = terms, 
                    sep = ":",
                    into = colmn,
                    remove = F) %>% 
    mutate(across(where(is.character), as.factor)) %>% 
    filter(significance == TRUE) %>% 

ggplot2::ggplot(data = .,
               # Highlight significant terms
               aes(x = terms, y = estimate, color = significance, shape = col_2)) +


            ggplot2::geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) +
            ggplot2::geom_point(size = 6, position = position_dodge(width = .9)) +

            # 95% C.I
            ggplot2::geom_linerange(aes(ymin = lower_bound,ymax = upper_bound),
                                    lwd = 1, position = position_dodge(width = .9)) +

            # 99% C.I
            #geom_linerange(aes(ymin = !!yvar - !!se * interval2,
            #                   ymax = !!yvar + !!se * interval2),
            #               lwd = 1/2, position = position_dodge(width = .9)) +

            ggplot2::theme_bw() +

            ggplot2::theme(legend.position = "none",
                           strip.text.x = element_text(size = 25),
                           axis.text.y   = element_text(size= 25),
                           axis.text.x   = element_text(size= 25),
                           axis.title.y  = element_text(size= 25),
                           axis.title.x  = element_text(size= 25),
                           panel.grid.major.y = element_blank(),
                           panel.grid.minor = element_blank(),
                           axis.line = element_line(size = .4,colour = "black"),
                           panel.border = element_rect(colour = "black", fill= NA,
                                                       size = 1.3)) +
            # Add name
            #labs(title = paste0(response_variable)) +

            # Significance colors
            ggplot2::scale_colour_manual(values = c("black","black")) +
            ggplot2::ylab("Estimated coefficients (median +/- 95CI)") +
            ggplot2::xlab("") +

            ggplot2::facet_wrap(~ response_var, scales = "free", ncol = 2) +


            ggplot2::coord_flip()

```

