---
title: 'Model fitting and validation'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Include code?
                      echo           = TRUE,
                      
                      error          = FALSE,
                      fig.align      = "center",
                      
                      # Path to where to store pdf single figures 
                      fig.path       = paste0("./figures_models_validation", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# # save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
# 
# for (file in doc.files) {
#   cambiar nombre
#     file.rename(file, file.path("../../hw1/", file))
# }
```


```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(dplyr)
# For Model fitting
library(lme4)
library(nlme)
library(purrr)
# For diagnostics
library(performance)
# For adding new columns
library(tibble)
```

```{r message=FALSE, warning=FALSE}
# Load data
sys.source("./scripts/code_join_data_full_dataset.R", envir = knitr::knit_global())
```


```{r message=FALSE, warning=FALSE}
# Load functions
## Models
sys.source("./R/functions_models.R", envir = knitr::knit_global())
```

# Select performance and traits variables 

## Transform variables
[Backtranformation info](http://www.biostathandbook.com/transformation.html#:~:text=Square%2Droot%20transformation.,to%20make%20them%20all%20positive.)

```{r}
data_for_models <-
    data_for_models %>%

        # Select variables for analysis
        dplyr::select(c(1:7,9,11:ncol(data_for_models)))
```


# Models: Questions 1 and 2   

$$response\sim treatment*fixer\ + initial\ height + random( 1|\ specie)$$

```{r}
# Take response variables names 
response_vars_q1_q2 <- 
    set_names(names(data_for_models)[c(5:12,14)])

response_vars_q1_q2 
```

```{r}
models_q1_q2 <- map(response_vars_q1_q2, ~ mixed_model_1(response = .x, 
                                                              data = data_for_models))
```

```{r}
model_q2_wue_log <- lmer(log(wue) ~ nfixer*treatment + init_height + (1|spcode), 
                         data = data_for_models)

model_q2_pnue_log <- lmer(log(pnue) ~ nfixer*treatment + init_height + (1|spcode), 
                         data = data_for_models) 

log_models <- list(model_q2_wue_log, model_q2_pnue_log)
names(log_models) <- c("wue_log", "pnue_log")
```

```{r}
# Append log models to model list
models_q1_q2 <- append(log_models, models_q1_q2)
```


## Models Nodule count

+ Chapter 9 Mixed models in ecology check glmmML package for count data
+ GOOD ref https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/
+ #https://www.flutterbys.com.au/stats/tut/tut11.2a.html

```{r}
# Load data
# This step was done like this because I am working with a subset of the data
# source cleaned data
source("./scripts/code_clean_data_nodules.R")

# Delete unused variables
data_nodules_cleaned <-
    data_nodules_cleaned %>%
        
        # add id to rownames for keep track of the rows
        column_to_rownames("id") %>% 
        dplyr::select(spcode,treatment, everything())

```

### m4 lmer gaussian      

```{r}
lmer_gaussian <- lmer(number_of_root_nodulation ~ treatment + init_height + 
                           (1 |spcode),data = data_nodules_cleaned)
```



```{r}
lmer_gaussian_log <-  lmer(log(number_of_root_nodulation) ~ treatment + init_height + 
                           (1 |spcode),data = data_nodules_cleaned)
```

### m5 glmer poisson

```{r}
glmer_poisson <-  glmer(number_of_root_nodulation ~ treatment + init_height + 
                           (1|spcode), family = "poisson",data = data_nodules_cleaned)
```


<!-- ## gls poisson  -->

<!-- ```{r} -->
<!-- gls_pois <- glm(number_of_root_nodulation ~ treatment*spcode + init_height,  -->
<!--                 data = data_nodules_cleaned,  -->
<!--                 family = poisson()) -->
<!-- ``` -->


<!-- ## linear model -->

<!-- ```{r} -->
<!-- lm_gaussian <-  lm(number_of_root_nodulation ~ treatment*spcode + init_height,  -->
<!--                      data = data_nodules_cleaned) -->

<!-- ``` -->



```{r}
models_nodule_count <- list(lmer_gaussian, lmer_gaussian_log, glmer_poisson)

names(models_nodule_count) <- c("lmer_gaussian","lmer_gaussian_log",
                                "glmer_poisson")

```

## Mycorrhizal colonization

```{r, eval = FALSE}
I decided not to include it, because I want to focus on Nfixing vrs non-Fixing, 
also I don't trust the data
```

# Models: Question 3


$$performance\sim treatment:fixer:scale(trait)\ + initial\ height + random( 1|\ specie)$$

## Scale preditors 

```{r}

data_for_models_scaled <-
    data_for_models %>% 
        mutate(across(c(4,9:14),scale)) 
```


```{r}
# test_again <- lme(total_biomass ~   treatment:nfixer:amax +
#                                     treatment:nfixer:gs  +
#                                     treatment:nfixer:wue +
#                                     treatment:nfixer:pnue +
#                                     treatment:nfixer:d13c +
#                                     init_height ,
#                   random = ~1 | spcode, 
#                   weights = varIdent(form = ~1 |spcode), 
#                   data = data_for_models_scaled)
# 
# 
# bootstrap_ci_results <- bootstrap_model_ci_df(model = test_again, 
#                                               category = "all", 
#                                               iter = 9999)

```




```{r}
# Select traits (x_vars)
traits_names <- 
    set_names(names(data_for_models_scaled)[c(9:12,14)])

traits_names
```

```{r}
# Select plants performance vars (y_vars)
performance_names <- 
    set_names(names(data_for_models_scaled)[c(5:8)])

performance_names
```

## Models lme4::lmer

```{r}
models_lmer_formulas <- model_combinations_formulas(performance_names, traits_names)

length(models_lmer_formulas)
models_lmer_formulas[1]
```

```{r}
models_q3_lmer <- map(models_lmer_formulas, 
                       ~ lmer(.x, data = data_for_models_scaled))
```

## Models nlme::lme

```{r}
models_nlme_formulas <- model_combinations_formulas(performance_names, 
                                                    traits_names, 
                                                    nlme = T)

length(models_nlme_formulas)
models_nlme_formulas[1]
```



```{r}
models_q3_nlme <- map(models_nlme_formulas, ~model_nlme(.,data = data_for_models_scaled))
```



# Validation plots 

# Collinearity

```{r collinearity_q1_q2}
map(models_q1_q2, check_collinearity)
```

```{r}
# map(models_nodule_count, check_collinearity)
```

```{r collinearity_q3}

#Warning: Model has interaction terms. VIFs might be inflated. You may check 
#multicollinearity among predictors of a model without interaction terms.

#map(models_list_q3, check_mul)
```

# Bolker's plots

```{r}
bolker_validation <- function(model) {
    
    
    a <- plot(model, type = c("p", "smooth"))
    
    ## heteroscedasticity
    b <-  plot(model,sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"))
   
    cowplot::plot_grid(a,b,nrow = 1)
}
```

## Models for questions 1 and 2 

```{r bolker_plots_q1_q2,fig.height = 8, fig.width = 15}
map(models_q1_q2, bolker_validation)
```

## Models for nodule count

```{r,bolker_plots_nodule_count, fig.height = 8, fig.width = 15}
map(models_nodule_count, plot)
```

## Models for question 3

### lme4 models

```{r,bolker_plots_q3, fig.height = 8, fig.width = 15}
map(models_q3_lmer, bolker_validation)
```

### nlme models

```{r,bolker_plots_q3_nlme, fig.height = 8, fig.width = 15}
map(models_q3_nlme, bolker_validation)
```

# Performance package

## Models for questions 1,2 

```{r validation_plots_q1_q2,fig.height = 15, fig.width = 10}
map(models_q1_q2, check_model)
```

## Models for nodule count

```{r validation_plots_nodule_count,fig.height = 15, fig.width = 10}
map(models_nodule_count, check_model)
```

## Models for question 3

```{r validation_plots_3_way_interact_lmer, fig.height = 15, fig.width = 10}
map(models_q3_lmer, check_model)
```



# Save lists with the models 

```{r}
saveRDS(models_q1_q2, file = "./processed_data/models_q1_q2.RData") 
saveRDS(models_q3_lmer, file = "./processed_data/models_q3_3_way_interaction_lmer.RData") 
saveRDS(models_q3_nlme, file = "./processed_data/models_q3_3_way_interaction_nlme.RData")
saveRDS(models_nodule_count, file = "./processed_data/models_nodule_count.RData") 
```
