---
title: 'Model fitting and validation'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Include code?
                      echo           = TRUE,
                      
                      error          = FALSE,
                      fig.align      = "center",
                      
                      # Path to where to store pdf single figures 
                      fig.path       = paste0("./figures_model_validation_lmer", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# # save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
# 
# for (file in doc.files) {
#   cambiar nombre
#     file.rename(file, file.path("../../hw1/", file))
# }
```


```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(dplyr)
# For Model fitting
library(lme4)
library(purrr)
# For diagnostics
library(performance)
# For adding new columns
library(tibble)
library(caret)
```

```{r message=FALSE, warning=FALSE}
# Load data
sys.source("./scripts/code_join_data_full_dataset.R", envir = knitr::knit_global())
```


```{r message=FALSE, warning=FALSE}
# Load functions
## Models
sys.source("./R/functions_models.R", envir = knitr::knit_global())
```

# Select performance and traits variables 


## Transform variables
[Backtranformation info](http://www.biostathandbook.com/transformation.html#:~:text=Square%2Droot%20transformation.,to%20make%20them%20all%20positive.)

## Back transformation

Even though you've done a statistical test on a transformed variable, such as 
the log of fish abundance, it is not a good idea to report your means, standard 
errors, etc. in transformed units. A graph that showed that the mean of the log 
of fish per 75 meters of stream was 1.044 would not be very informative for 
someone who can't do fractional exponents in their head. Instead, you should 
back-transform your results. This involves doing the opposite of the mathematical 
function you used in the data transformation. For the log transformation, you 
would back-transform by raising 10 to the power of your number. For example, the 
log transformed data above has a mean of 1.044 and a 95% confidence 
interval of ±0.344 log-transformed fish. The back-transformed mean would 
be 101.044=11.1 fish. The upper confidence limit would be 10(1.044+0.344)=24.4 
fish, and the lower confidence limit would be 10(1.044-0.344)=5.0 fish. 
Note that the confidence interval is not symmetrical; the upper limit is 13.3 
fish above the mean, while the lower limit is 6.1 fish below the mean. Also note 
that you can't just back-transform the confidence interval and add or subtract 
that from the back-transformed mean; you can't take 100.344 and add or subtract that.


Square-root transformation. This consists of taking the square root of each 
observation. The back transformation is to square the number. If you have 
negative numbers, you can't take the square root; you should add a constant to 
each number to make them all positive.

People often use the square-root transformation when the variable is a count of 
something, such as bacterial colonies per petri dish, blood cells going through 
a capillary per minute, mutations per generation, etc.



```{r}
# Box-Cox transformation was previously run 

data_for_models_transformed <- 
    data_for_models %>% 
        
        mutate(
            # Plant's performance
            #total_biomass_sqrt = sqrt(total_biomass),
            #above_biomass_sqrt = sqrt(above_biomass),
            #below_biomass_log = log(below_biomass),
            #agr_log = log(agr),
            
            # NO TRANSFORMATION variable already in log-log
            #rgr = rgr,
            
            # NO TRANSFORMATION variable already in log-log
            #rgr_slope = rgr_slope,
          
            # Traits
            amax_log = log(amax),
            gs_log = log(gs),
            wue_log = log(wue),
            
            # d13 and d15 where not transformed because the data has negative values
            pnue_log = log(data_for_models$pnue),
          
            # Covariate
            init_height = log(init_height)) %>%   
            
        # Remove original variables (non-transformed)
        dplyr::select(spcode, treatment, nfixer, init_height, everything(),
               -c(8,10,11:13,16))    

```



# Models: Questions 1 and 2   

$$response\sim treatment*fixer\ + initial\ height + random( 1|\ specie)$$

```{r}
# Take response variables' names 
response_vars_q1_q2 <- 
    set_names(names(data_for_models_transformed)[5:(ncol(data_for_models_transformed))])
```



```{r}
#data_for_models_transformed[data_for_models_transformed$id == 48,]
models_list_q1_q2 <- map(response_vars_q1_q2, ~ mixed_model_1(response = .x, 
                                                              data = data_for_models_transformed))
```

## Models Nodule count

+ Chapter 9 Mixed models in ecology check glmmML package for count data
+ GOOD ref https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/
+ #https://www.flutterbys.com.au/stats/tut/tut11.2a.html

```{r}
# Load data
# This step was done like this because I am working with a subset of the data
# source cleaned data
source("./scripts/code_clean_data_nodules.R")

# Delete unused variables
data_nodules_cleaned <-
    data_nodules_cleaned %>%
        
        # add id to rownames for keep track of the rows
        column_to_rownames("id") %>% 
        dplyr::select(spcode,treatment, everything())

```


### m4 lmer gaussian      

```{r}
lmer_gaussian <- lmer(number_of_root_nodulation ~ treatment + init_height + 
                           (1 |spcode),
                       data = data_nodules_cleaned)
```


```{r}
lmer_gaussian_log <-  lmer(log(number_of_root_nodulation) ~ treatment + init_height + 
                           (1 |spcode),
                       data = data_nodules_cleaned)
```

### m5 glmer poisson

```{r}
glmer_poisson <-  glmer(number_of_root_nodulation ~ treatment + init_height + 
                           (1 |spcode), family = "poisson",
                       data = data_nodules_cleaned)
```

```{r}
models_list_nodule_count <- list(lmer_gaussian, lmer_gaussian_log, glmer_poisson)
names(models_list_nodule_count) <- c("lmer_gaussian",
                                     "lmer_gaussian_log",
                                     "glmer_poisson")
```


## Mycorrhizal colonization

```{r, eval = FALSE}
I decided not to include it, because I want to focus on Nfixing vrs non-Fixing, 
also I don't trust on the data
```


# Models: Question 3

$$performance\sim treatment*\ fixer*\ scale(log(trait)\ + initial\ height + random( 1|\ specie)$$

## Scale preditors

```{r}

data_for_models_transformed_scaled <-         
    data_for_models_transformed %>% 
    
    # test for being sure that I select the traits
    #select(c(4,7,8,13:16))

        # scale only the predictors traits
         mutate(across(c(9:14), scale))
```


```{r}
# Select traits (x_vars)
traits_names <- 
    set_names(names(data_for_models_transformed_scaled)
                          [c(9:14)])
traits_names
```


```{r}
# Select plants performance vars (y_vars)
performance_names <- 
    set_names(names(data_for_models_transformed_scaled)[c(5:8)])

performance_names
```


```{r}
models_lmer_formulas <- model_combinations_formulas(performance_names, traits_names)

length(models_lmer_formulas)
models_lmer_formulas[1]
```

```{r}
models_list_q3 <- map(models_lmer_formulas, 
                       ~ lmer(.x, data = data_for_models_transformed_scaled))
```

# Validation plots 

# Collinearity

```{r collinearity_q1_q2}
map(models_list_q1_q2, check_collinearity)
```

```{r}
map(models_list_nodule_count, check_collinearity)
```

```{r collinearity_q3}

#Warning: Model has interaction terms. VIFs might be inflated. You may check 
#multicollinearity among predictors of a model without interaction terms.

#map(models_list_q3, check_mul)
```


# Bolker's plots

```{r}
bolker_validation <- function(model) {
    
    
    a <- plot(model, type = c("p", "smooth"))
    
    ## heteroscedasticity
    b <-     plot(model,sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"))
   
    cowplot::plot_grid(a,b,nrow = 1)
}
```

## Models for questions 1,2 

```{r bolker_plots_q1_q2,fig.height = 8, fig.width = 15}
map(models_list_q1_q2, bolker_validation)
```

## Models for nodule count

```{r, bolker_plots_q3, fig.height = 8, fig.width = 15}
map(models_list_nodule_count, bolker_validation)
```

## Models for question 3
```{r, bolker_plots_nodule_count, fig.height = 8, fig.width = 15}
map(models_list_q3, bolker_validation)
```


# Performance package

## Models for questions 1,2 
```{r validation_plots_q1_q2,fig.height = 15, fig.width = 10}
map(models_list_q1_q2, check_model)
```


## Models for nodule count
```{r validation_plots_nodule_count,fig.height = 15, fig.width = 10}
map(models_list_nodule_count, check_model)
```


## Models for question 3
```{r validation_plots_3_way_interact,fig.height = 15, fig.width = 10}
map(models_list_q3, check_model)
```



# Save lists with the models 

```{r}
saveRDS(models_list_q1_q2, file = "./processed_data/models_q1_q2.RData") 
saveRDS(models_list_q3, file = "./processed_data/models_q3_3_way_interaction.RData") 
saveRDS(models_list_nodule_count, file = "./processed_data/models_list_nodule_count.RData") 
```
