---
title: 'Model fitting and validation'
author: "Erick Calderon-Morales"
date: ' Fall 2021'
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
                    fig.width = 11, fig.height = 7)
```


```{r knitr, include = FALSE}

# Save figures in specific place
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,

                      # Include code?
                      echo           = TRUE,

                      error          = FALSE,
                      fig.align      = "center",

                      # Path to where to store pdf single figures
                      fig.path       = paste0("./figures_models_validation", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r cleanup-docs, cache = FALSE,echo = FALSE}

# # save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#   cambiar nombre
#     file.rename(file, file.path("../../hw1/", file))
# }
```


```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(dplyr)
# For Model fitting
library(lme4)
library(nlme)
library(purrr)
# For diagnostics
library(performance)
#library(see)
# For adding new columns
library(tibble)
library(MuMIn)
```
```{r}
# setwd for Rmarkdowm
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r message=FALSE, warning=FALSE}
# Load data
source("./scripts/code_join_data_full_dataset.R")
```


```{r message=FALSE, warning=FALSE}
# Load functions
## Models
source("./R/functions_models.R")
source("./R/function_nlme_validation_plots.R")
```

# Model Fitting

## Models: Questions 1,2 and Mass fractions

### Select performance and traits variables

```{r}
data_for_models <-
    data_for_models %>%

        # Select variables for analysis
        dplyr::select(!c(rgr_slope, d15n, above_biomass, below_biomass,
                        rmf, smf, lmf))
```



$$response\sim treatment*fixer\ + initial\ height + random( 1|\ specie)$$

```{r}
# Take response variables names
response_vars_q1_q2 <-
    set_names(names(data_for_models)[c(5:14)])

response_vars_q1_q2
```

```{r}
models_q1_q2 <- map(response_vars_q1_q2, ~ mixed_model_1(response = .x,
                                                        data = data_for_models))

names(models_q1_q2)
```

```{r}
# Log models

## WUE log
model_q2_wue_log <- lmer(log(wue) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)
## PNUE log
model_q2_pnue_log <- lmer(log(pnue) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

## Root-Shoot ratio
model_ratio_log <- lmer(log(root_shoot_ratio) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)
## Nitrogen log
model_nitrogen_log <- lmer(log(Narea_g_m2) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

## Absolute growth rate
model_agr_log <- lmer(log(agr) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

```


```{r}
# Append log models to model list Q1 Q2
log_models <- list(model_q2_wue_log, model_q2_pnue_log,
                   model_ratio_log, model_nitrogen_log,
                   model_agr_log)


names(log_models) <- c("wue_log", "pnue_log", "ratio_log", "nitrogen_log",
                        "agr_log")

models_q1_q2 <- append(log_models, models_q1_q2)

length(models_q1_q2)
names(models_q1_q2)
```

## Model: Nodule count

+ Chapter 9 Mixed models in ecology check glmmML package for count data
+ GOOD ref https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/
+ #https://www.flutterbys.com.au/stats/tut/tut11.2a.html

```{r}
# Load data
# This step was done like this because I am working with a subset of the data
# source cleaned data
source("./scripts/code_clean_data_nodules.R")

# Delete unused variables
data_nodules_cleaned <-
    data_nodules_cleaned %>%

        # add id to rownames for keep track of the rows
        column_to_rownames("id") %>%
        dplyr::select(spcode, treatment, everything())

```


__Previuosly I fitted non-log models and found out the log one is the best__

### log models

```{r}
## After seeing the validation plots for the lmer_gaussian_log I decided to try
## to improve the log model using the nlme package

nlme_gaussian_log <- lme(log(number_of_root_nodulation) ~ treatment + init_height,
                                    random = ~1|spcode,
                                    data = data_nodules_cleaned)


nlme_nodule_log_weights <- lme(log(number_of_root_nodulation) ~ treatment + init_height,
                                    random = ~1|spcode,
                                    weights = varIdent(form = ~1|spcode),
                                    data = data_nodules_cleaned)
```

#### Compare models
```{r}
# Check if the nlme_gaussian_log_weights is a better model

# Compare between nlme objects
anova(nlme_gaussian_log, nlme_nodule_log_weights)

model.sel(nlme_gaussian_log, nlme_nodule_log_weights)

AIC(nlme_gaussian_log, nlme_nodule_log_weights)
# Model improved, check assumptions for nlme_gaussian_log_weights

```


## Models: Question 3

$$performance\sim treatment:fixer:scaled(trait)\ + initial\ height + random( 1|\ specie)$$

### Scale preditors aka traits

```{r}
data_for_models_scaled <-
    data_for_models %>%

    #  Narea_g_m2 not included since it has a high correlation with amax
    mutate(across(c(4, 9:13), scale))
```



### Model RGR

```{r}
lme_rgr_3way <- lme(rgr ~   treatment:nfixer:amax +
                            treatment:nfixer:gs  +
                            treatment:nfixer:wue +
                            treatment:nfixer:pnue +
                            treatment:nfixer:d13c +

                            # Added
                            #treatment:nfixer:Narea_g_m2 +
                            init_height ,
                  random = ~1 | spcode,
                  data = data_for_models_scaled)
```


### Model AGR

```{r}
lme_agr_3way <- lme(agr ~   treatment:nfixer:amax +
                            treatment:nfixer:gs  +
                            treatment:nfixer:wue +
                            treatment:nfixer:pnue +
                            treatment:nfixer:d13c +

                            # Added
                            #treatment:nfixer:Narea_g_m2 +
                            init_height ,
                  random = ~1 | spcode,
                  data = data_for_models_scaled)
```


### Model Total Biomass

```{r}
lme_total_biom_3way <- lme(total_biomass ~  treatment:nfixer:amax +
                                            treatment:nfixer:gs  +
                                            treatment:nfixer:wue +
                                            treatment:nfixer:pnue +
                                            treatment:nfixer:d13c +

                                            init_height,
                            random = ~1 | spcode,
                            data = data_for_models_scaled)
```

## Model shoot root ratios

```{r}
lme_3way_root_shoot_ratio <- lme(root_shoot_ratio ~  treatment:nfixer:amax +
                                                        treatment:nfixer:gs  +
                                                        treatment:nfixer:wue +
                                                        treatment:nfixer:pnue +
                                                        treatment:nfixer:d13c +

                                                        # Added
                                                        #treatment:nfixer:Narea_g_m2 +
                                                        init_height,
                            random = ~1 | spcode,
                            data = data_for_models_scaled)
```

# Model Validation


## Validation of models for questions 1,2 and Mass fractions

### Collinearity

```{r collinearity_q1_q2}
map(models_q1_q2, check_collinearity)
```


```{r validation_plots_q1_q2,fig.height = 15, fig.width = 10}
# No problems found here
map(models_q1_q2, check_model)
```


## Validation of nodule count models

### Checking log nodule count model's assumptions


Zuur et al pp 84:

"Note that these residuals still show heterogeneity, but this is now allowed
(because the residual variation differs depending on the chosen variance
structure and values of the variance covariate). Hence, these residuals are
less useful for the model validation process."

```{r, validation_plots_nodule_log_weights,  fig.height = 9, fig.width = 14}
nlme_validation_plots(nlme_nodule_log_weights, data = data_nodules_cleaned,
                      group = "spcode", variables = c("treatment") )
```

## Validation of models for question 3

### RGR

```{r validation_rgr_3way, fig.height = 9, fig.width = 14}
nlme_validation_plots(lme_rgr_3way, data = data_for_models_scaled, group = "spcode")
```

### AGR
```{r validation_agr_3way, fig.height = 9, fig.width = 14}
nlme_validation_plots(lme_agr_3way, data = data_for_models_scaled, group = "spcode")
```


### Total biomass

```{r validation_lme_total_biom_3way, fig.height = 9, fig.width = 14}
nlme_validation_plots(lme_total_biom_3way, data = data_for_models_scaled,
                      group = "spcode")
```

### Root-shoot ratio

```{r}
nlme_validation_plots(lme_3way_root_shoot_ratio, data = data_for_models_scaled,
                      group = "spcode")
```


# Save model lists as .RData

# Models for Q1, Q2 and Mass fractions

```{r}
models_q1_q2 <-

    models_q1_q2 %>%

    # Remove models that showed violation of the assumptions
    purrr::list_modify("pnue" = NULL, "wue" = NULL,
                       "Narea_g_m2" = NULL, "root_shoot_ratio" = NULL)

```

## Nodule count

```{r}
models_nodule_count <- list(nlme_gaussian_log, nlme_nodule_log_weights)
names(models_nodule_count) <- c("nlme_gaussian_log", "nlme_gaussian_log_weights")
```

## 3-way models

```{r}
models_3way_q3 <- list(lme_rgr_3way, lme_total_biom_3way,
                            lme_3way_root_shoot_ratio)

names(models_3way_q3) <- c("lme_rgr_3way", "lme_total_biom_3way",
                                "lme_3way_root_shoot_ratio")
```


```{r}
saveRDS(models_q1_q2, file = "./processed_data/models_q1_q2.RData")
saveRDS(models_nodule_count, file = "./processed_data/models_nodule_count.RData")
saveRDS(models_3way_q3, file = "./processed_data/models_3way_q3.RData")
```