---
title: 'Model fitting and validation for Q2'
author: "Erick Calderon-Morales"
date: ''
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: tactile
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r, setup, include=FALSE}
# Set knit directory
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(
    autodep = TRUE,
    cache = FALSE,
    cache.comments = TRUE,

  # Save figures as pdf ?
  # dev = c( "png", "pdf"),

  # Include code?
  echo = TRUE,
  error = FALSE,
  fig.align = "center",

  # Path where figures are going to be store pdf single
  # figures
  fig.path = paste0("figures/"),
  fig.width = 11,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

```{r cleanup-docs, cache = FALSE,echo = FALSE}
# save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#     file.rename(file, file.path("../../[insert_folder_name]/", file))
# }
```


```{r libaries, message = FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For Model fitting
library(lme4)
library(nlme)
library(purrr)
# For diagnostics
library(performance)
# library(see)
# For adding new columns
library(tibble)
library(MuMIn)
library(ggplot2)
```

```{r}
# setwd for Rmarkdowm
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r message=FALSE, warning=FALSE}
# Load data
source("./scripts/code_join_data_full_dataset.R")
```


```{r message=FALSE, warning=FALSE}
# Load functions
## Models
source("./R/functions_models.R")
source("./R/function_nlme_validation_plots.R")

## Plots
source("./R/function_masomenos_plots.R")
```

# Model Fitting

__Q2: How does increased nutrient and/or water availability influence seedling
water and nutrient-use traits and the relationships with N-fixing bacteria?__

```{r}
data_for_models <-
    data_for_models %>%

    # Select variables for analysis
    dplyr::select(!c(rgr_slope, d15n, above_biomass, below_biomass, rmf, smf,
                    lmf, agr, d13c ))
```

## Model formula

$$response\sim treatment*fixer\ + initial\ height + random( 1|\ specie)$$

## Select response variables

```{r}
# Take response variables names
response_vars_q2 <-
  set_names(names(data_for_models)[c(8:12)])

response_vars_q2
```


## Fit models

###  Traits

```{r}
models_q2 <- map(response_vars_q2, ~ mixed_model_1(response = .x,
                    data = data_for_models))

names(models_q2)
```


### Nodule count

+ Chapter 9 Mixed models in ecology check glmmML package for count data
+ GOOD ref https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/
+ #https://www.flutterbys.com.au/stats/tut/tut11.2a.html


```{r}
# Load data
# This step was done like this because I am working with a subset of the data
# source cleaned data
source("./scripts/code_clean_data_nodules.R")

# Delete unused variables
data_nodules_cleaned <-
    data_nodules_cleaned %>%

        # add id to rownames for keep track of the rows
        column_to_rownames("id") %>%
        dplyr::select(spcode, treatment, everything())

```


```{r}
#https://www.dataquest.io/blog/tutorial-poisson-regression-in-r/
lm_nodule_treat_spcode <- lm(number_of_root_nodulation ~ treatment * spcode +
                                                            init_height,
                                data = data_nodules_cleaned)
```


```{r}
#__Previuosly I fitted non-log models and found out the log one is the best__

## After seeing the validation plots for the lmer_gaussian_log I decided to try
## to improve the log model using the nlme package

#nlme_gaussian_log <- lme(log(number_of_root_nodulation) ~ treatment + init_height,
#                                    random = ~1|spcode,
#                                    data = data_nodules_cleaned)


#nlme_nodule_log_weights <- lme(log(number_of_root_nodulation) ~ treatment + init_height,
#                                    random = ~1|spcode,
#                                    weights = varIdent(form = ~1|spcode),
#                                    data = data_nodules_cleaned)
```

#### Compare models
```{r}
# Check if the nlme_gaussian_log_weights is a better model

# Compare between nlme objects
#anova(nlme_gaussian_log, nlme_nodule_log_weights)

#model.sel(nlme_gaussian_log, nlme_nodule_log_weights)

#AIC(nlme_gaussian_log, nlme_nodule_log_weights)
# Model improved, check assumptions for nlme_gaussian_log_weights
```


# Model Validation

## Collinearity

```{r collinearity_q2}
map(models_q2, check_collinearity)
```

```{r}
check_collinearity(lm_nodule_treat_spcode)
```

## Assumptions

```{r validation_plots_q2, fig.height = 15, fig.width = 10}
# No problems found here
map(models_q2, check_model)
```


### Checking log nodule count model's assumptions

Zuur et al pp 84:

"Note that these residuals still show heterogeneity, but this is now allowed
(because the residual variation differs depending on the chosen variance
structure and values of the variance covariate). Hence, these residuals are
less useful for the model validation process."

```{r, validation_plots_nodule_log_weights,  fig.height = 9, fig.width = 14}
#nlme_validation_plots(nlme_nodule_log_weights, data = data_nodules_cleaned,
#                      group = "spcode", variables = c("treatment"))
```


```{r lm_nodule_model_assumptions}
check_model(lm_nodule_treat_spcode)
```

# Refit models that showed violation of any assumptions

```{r}
# Log models

## WUE log
model_q2_wue_log <- lmer(log(wue) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

## PNUE log
model_q2_pnue_log <- lmer(log(pnue) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

## Nitrogen log
model_q2_nitrogen_log <- lmer(log(Narea_g_m2) ~ nfixer*treatment + init_height + (1|spcode),
                         data = data_for_models)

```


```{r}
# Append log models to model list Q2
log_models_q2 <- list(model_q2_wue_log, model_q2_pnue_log,
                      model_q2_nitrogen_log)

names(log_models_q2) <- c("wue_log", "pnue_log", "nitrogen_log")
```


# Check assumptions again

```{r collinearity_q2_log}
map(log_models_q2, check_collinearity)
```


```{r validation_plots_q2_log, fig.height = 15, fig.width = 10}
# No problems found here
map(log_models_q2, check_model)
```


```{r}
# Remove models that showed bad performance
models_q2 <-

    models_q2 %>%

    # Remove models that showed violation of the assumptions
    purrr::list_modify("pnue" = NULL, "wue" = NULL, "Narea_g_m2" = NULL)
```

```{r}
# Append log models
models_q2 <- append(log_models_q2, models_q2)

length(models_q2)
names(models_q2)
```


# Save model lists as .RData

```{r}
saveRDS(models_q2, file = "./processed_data/models_q2.RData")

# Save nodule count model
saveRDS(lm_nodule_treat_spcode, file = "./processed_data/model_nodule_interaction.RData")
```