---
title: 'Model fitting and validation for Q3'
author: "Erick Calderon-Morales"
date: ''
due_date: ""
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: tactile
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r, setup, include=FALSE}
# Set knit directory
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(
  autodep = TRUE,
  cache = FALSE,
  cache.comments = TRUE,

  # Save figures as pdf ?
  # dev = c( "png", "pdf"),

  # Include code?
  echo = TRUE,
  error = FALSE,
  fig.align = "center",

  # Path where figures are going to be store pdf single
  # figures
  fig.path = paste0("figures/"),
  fig.width = 11,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

```{r cleanup-docs, cache = FALSE,echo = FALSE}
# save a html copy file in a specific place
# doc.files <- c(list.files(pattern = "pdf"),
#                list.files(pattern = "html"),
#                list.files(pattern = "docx"))
#
# for (file in doc.files) {
#     file.rename(file, file.path("../../[insert_folder_name]/", file))
# }
```


```{r libaries, message = FALSE, warning = FALSE, cache = FALSE}
library(dplyr)
# For Model fitting
library(lme4)
library(nlme)
library(purrr)
# For diagnostics
library(performance)
# library(see)
# For adding new columns
library(tibble)
library(MuMIn)
```

```{r}
# setwd for Rmarkdowm
knitr::opts_knit$set(root.dir = "~/Documents/projects/nutrients_and_water_effects_2022/stats/")
```


```{r message=FALSE, warning=FALSE}
# Load data
source("./scripts/code_join_data_full_dataset.R")
```


```{r message=FALSE, warning=FALSE}
# Load functions
## Models
source("./R/functions_models.R")
source("./R/function_nlme_validation_plots.R")
```



__Q3: How do seedling-level functional traits, treatment, and N-fixing status
interact to influence seedling-level growth rates and biomass accumulation?__

```{r}
data_for_models <-
  data_for_models %>%
  # Select variables for analysis
  dplyr::select(!c(rgr_slope, d15n, above_biomass, below_biomass, amax,
                  rmf, smf, lmf, agr))
```


### Scale preditors (traits)

```{r}
data_for_models_scaled <-
    data_for_models %>%
    mutate(across(
    c("init_height", "gs", "wue", "pnue", "Narea_g_m2",
      "rgr", "root_shoot_ratio", "total_biomass"), scale))
```


# Model formula

$$performance\sim treatment:fixer:scaled(trait)\ + initial\ height + random( 1|\ specie)$$


# Model RGR

```{r}
lme_rgr_3way <- lme(rgr ~ treatment:nfixer:Narea_g_m2 +
                          treatment:nfixer:gs +
                          treatment:nfixer:wue +
                          treatment:nfixer:pnue +

                        init_height,
                        random = ~ 1 | spcode,
                        data = data_for_models_scaled)
```

## Model Validation

```{r validation_rgr_3way, fig.height = 12, fig.width = 15}
nlme_validation_plots(lme_rgr_3way, data = data_for_models_scaled, group = "spcode")
```


# Model shoot root ratios

```{r}
lme_root_shoot_3way <- lme(root_shoot_ratio ~ treatment:nfixer:Narea_g_m2 +
                                              treatment:nfixer:gs +
                                              treatment:nfixer:wue +
                                              treatment:nfixer:pnue +

                                              init_height,
                                              random = ~ 1 | spcode,
                                          data = data_for_models_scaled)
```


## Model Validation

```{r validation_root_shoot_3way, fig.height = 12, fig.width = 15}
nlme_validation_plots(lme_root_shoot_3way, data = data_for_models_scaled,
                      group = "spcode")
```

# Model Total Biomass

```{r}
lme_total_biom_3way <- lme(total_biomass ~ treatment:nfixer:Narea_g_m2 +
                                            treatment:nfixer:gs +
                                            treatment:nfixer:wue +
                                            treatment:nfixer:pnue +

                                            init_height,
                                            random = ~ 1 | spcode,
                                            data = data_for_models_scaled)
```

## Model Validation

```{r validation_lme_total_biom_3way, fig.height = 12, fig.width = 15}
nlme_validation_plots(lme_total_biom_3way, data = data_for_models_scaled,
                      group = "spcode")
```


# Save model lists as .RData

## 3-way models

```{r}
models_3way_q3 <- list(lme_rgr_3way, lme_total_biom_3way, lme_root_shoot_3way)

names(models_3way_q3) <- c("lme_rgr_3way", "lme_total_biom_3way",
                            "lme_3way_root_shoot_ratio")
```


# lme4 models for double check

## rgr
```{r}
lmer_rgr_3way <- lmer(rgr ~   treatment:nfixer:Narea_g_m2 +
                              treatment:nfixer:gs  +
                              treatment:nfixer:wue +
                              treatment:nfixer:pnue +

                              (1 | spcode) +
                              init_height,
                  data = data_for_models_scaled)
```

```{r, fig.height = 15, fig.width = 12}
performance::check_model(lmer_rgr_3way)
```

## Root shoot ratio
```{r}
lmer_root_shoot_3way <- lmer(root_shoot_ratio ~   treatment:nfixer:Narea_g_m2 +
                                                  treatment:nfixer:gs  +
                                                  treatment:nfixer:wue +
                                                  treatment:nfixer:pnue +

                              (1 | spcode) +
                              init_height,
                  data = data_for_models_scaled)
```

```{r, fig.height = 15, fig.width = 12}
performance::check_model(lmer_root_shoot_3way)
```


## Total biomass

```{r}
lmer_total_biom_3way <- lmer(total_biomass ~ treatment:nfixer:Narea_g_m2 +
                                              treatment:nfixer:gs +
                                              treatment:nfixer:wue +
                                              treatment:nfixer:pnue +

                                              init_height + (1 | spcode),
                              data = data_for_models_scaled)
```

```{r, fig.height = 15, fig.width = 12}
performance::check_model(lmer_total_biom_3way)
```